---
title: "Week45 - shotsignR"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set( 
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## Some Love for `queryBuildR`

Somehow I missed [`queryBuildR`](https://github.com/Yannael/queryBuildR) when it was released.  I strongly recommend checking it out.

## This Week's Widget - shotsignR

[Peter Beshai](http://peterbeshai.com) created beautiful `d3.js` [Shooting Signatures](https://gist.github.com/pbeshai/ffd0f9d84b4e8df27db2) to visualize basketball shooting performance.  It was definitely love at first sight for me.  However, to justify spending time creating an `htmlwidget`, I needed some inspiration and also a non-basketball use case.  Both the inspiration and use case came from [Kevin Ferris \@kferris10](https://github.com/kferris10) in this [issue](https://github.com/timelyportfolio/buildingwidgets/issues/14).

Unfortunately, this is still very much a work-in-progress.  I'm hoping for some additional inspiration/help to finish it up.  Also, midway through this project, I realized that this is sort-of a modified `streamgraph`, so I might circle back to see if I can build on top of Bob Rudis' excellent [streamgraph package](https://github.com/hrbrmstr/streamgraph).

## Installation

This is not on CRAN, so to install we will need some help from `devtools::install_github`.

```{r eval = FALSE}
devtools::install_github("timelyportfolio/shotsignR")
```

## Examples

### Basketball Example

```{r "example01", eval = F}
# devtools::install_github("timelyportfolio/shotsignR")

library("shotsignR")

# use data from Peter Beshai gist https://gist.github.com/pbeshai/ffd0f9d84b4e8df27db2
test_data <- jsonlite::fromJSON(
  paste0(
    'https://gist.githubusercontent.com/',
    'timelyportfolio/d00488979c22fc4ed900/raw/',
    'a7838e5ea1abe9f0d836e55499bd618efaeae5b7/shotsign.json'
  )
)

shotsign( test_data )
```

```{r "example01_run", echo = F}
# devtools::install_github("timelyportfolio/shotsignR")

library("shotsignR")

# use data from Peter Beshai gist https://gist.github.com/pbeshai/ffd0f9d84b4e8df27db2
test_data <- jsonlite::fromJSON(
  paste0(
    'https://gist.githubusercontent.com/',
    'timelyportfolio/d00488979c22fc4ed900/raw/',
    'a7838e5ea1abe9f0d836e55499bd618efaeae5b7/shotsign.json'
  )
)

shotsign( test_data, width = 600, height = 200 ) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 750
      , height = 350
      , overflow = "hidden"
    )
  )
```

### Non-basketball Usage

```{r "example02", eval = F}
# devtools::install_github("timelyportfolio/shotsignR")

library("shotsignR")

# now let's try the example from @kferris10
library(dplyr)
library(magrittr)
library(ggplot2)

x <- read.table("http://www.stat.ufl.edu/~winner/data/ecophys_rc.dat") %>% 
  setNames(c("location", "treatment", "plant_id", "co2_concentration", "co2_uptake_rate")) %>% 
  group_by(plant_id) %>% 
  mutate(time = 1:n()) %>% 
  ungroup()

x %>% 
  filter(plant_id == 1) %>% 
  qplot(time, co2_uptake_rate, data = ., geom = "line", group = plant_id, 
        colour = co2_concentration, size = co2_concentration)

x %>% 
  filter(plant_id == 1) %>%
  select(time, co2_uptake_rate, co2_concentration, co2_concentration) %>%
  set_colnames(c("x","y","widthValue")) %>%
  mutate(colorValue = widthValue/max(widthValue)) %>%
  {
    shotsign(
      .,
      xdomain = range(.$x),
      ydomain = range(.$y),
      wdomain = range(.$w),
      colordomain = c(0,1)
    )
  }

```

```{r "example02_run", echo = F}
# devtools::install_github("timelyportfolio/shotsignR")

library("shotsignR")

# now let's try the example from @kferris10
library(dplyr)
library(magrittr)
library(ggplot2)

x <- read.table("http://www.stat.ufl.edu/~winner/data/ecophys_rc.dat") %>% 
  setNames(c("location", "treatment", "plant_id", "co2_concentration", "co2_uptake_rate")) %>% 
  group_by(plant_id) %>% 
  mutate(time = 1:n()) %>% 
  ungroup()

x %>% 
  filter(plant_id == 1) %>% 
  qplot(time, co2_uptake_rate, data = ., geom = "line", group = plant_id, 
        colour = co2_concentration, size = co2_concentration)

x %>% 
  filter(plant_id == 1) %>%
  select(time, co2_uptake_rate, co2_concentration, co2_concentration) %>%
  set_colnames(c("x","y","widthValue")) %>%
  mutate(colorValue = widthValue/max(widthValue)) %>%
  {
    shotsign(
      .,
      xdomain = range(.$x),
      ydomain = range(.$y),
      wdomain = range(.$w),
      colordomain = c(0,1)
    , width = 600, height = 200 )
  } %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example02.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 750
      , height = 350
      , overflow = "hidden"
    )
  )
```




## Thanks

Thanks [Peter Beshai](http://peterbeshai.com) for [Shooting Signatures](https://gist.github.com/pbeshai/ffd0f9d84b4e8df27db2).

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

