---
title: "Week44 - tooltipsterR"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set( 
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

save_tags <- function (tags, file, selfcontained = F, libdir = "./lib") 
{
    if (is.null(libdir)) {
        libdir <- paste(tools::file_path_sans_ext(basename(file)), 
            "_files", sep = "")
    }
    htmltools::save_html(tags, file = file, libdir = libdir)
    if (selfcontained) {
        if (!htmlwidgets:::pandoc_available()) {
            stop("Saving a widget with selfcontained = TRUE requires pandoc. For details see:\n", 
                "https://github.com/rstudio/rmarkdown/blob/master/PANDOC.md")
        }
        htmlwidgets:::pandoc_self_contained_html(file, file)
        unlink(libdir, recursive = TRUE)
    }
    return(htmltools::tags$iframe(src= file, height = "400px", width = "100%", style="border:0;"))
}


library(pipeR)

```

## Only 7 More Weeks

As I typed "Week 44" in the title of the post, I realized the year is almost done.  I do not anticipate another widget-a-week commitment next year, but I certainly will make some commitment.  Let me know if you have ideas.  Also, I have some `htmlwidgets` in the pipeline, but not seven, so please let me know if you have a strong compelling story for a `htmlwidget` to end the year with a bang.

## This Week's Widget - tooltipsterR

Tooltips are an invaluable but often neglected component of a visualization or story.  The jQuery plugin [`tooltipster`](http://iamceege.github.io/tooltipster/) is a mature, well-documented, and full-featured tooltip library, so I feel like we need it in `R` and thus [`tooltipsterR`](https://github.com/timelyportfolio/tooltipsterR).


## Installation

This is not on CRAN, so to install we will need some help from `devtools::install_github`.

```{r eval = FALSE}
devtools::install_github("timelyportfolio/tooltipsterR")
```

## Examples

Check out the nicely done [tooltipster demos](http://iamceege.github.io/tooltipster/#demos) to see all the power and beauty of `tooltipster`.  I'll just show some basic examples below.


### Easy

Just by adding `class="tooltip"` and `title=...`, we can get tooltips.

```{r "example01", eval = F}
#devtools::install_github("timelyportfolio/tooltipsterR")

library(tooltipsterR)
library(htmltools)

tagList(
  tooltipster(),
  tags$p(
    "See if you can find my ",
    tags$span(
      class = "tooltip", # can change if desired
      title = "you found me",
      style = "border-bottom: 1px dashed #999;",      
      "tooltip"
    )
  ),
  tags$div(
    class = "tooltip",
    title = "nowhere to hide",
    style = "height:200px;width:200px;",
    style = "border-radius:100%;background:#B062C3;",
    tags$span(
      style = "text-align:center;padding-top:5.5em;display:block;",
      "not here"
    )
  )
)
```

```{r "example01_run", echo = F}
#devtools::install_github("timelyportfolio/tooltipsterR")

library(tooltipsterR)
library(htmltools)

tagList(
  tooltipster(),
  tags$p(
    "See if you can find my ",
    tags$span(
      class = "tooltip", # can change if desired
      title = "you found me",
      style = "border-bottom: 1px dashed #999;",      
      "tooltip"
    )
  ),
  tags$div(
    class = "tooltip",
    title = "nowhere to hide",
    style = "height:200px;width:200px;",
    style = "border-radius:100%;background:#B062C3;",
    tags$span(
      style = "text-align:center;padding-top:5.5em;display:block;",
      "not here"
    )
  )
) %>>%
  save_tags(file="example01.html")
```


### With Formattable

For a little more advanced example, let's combine `tooltipsterR` with week 23's [`formattable`](http://www.buildingwidgets.com/blog/2015/6/12/week-23-formattable).

```{r "example02", eval = F}
#devtools::install_github("timelyportfolio/tooltipsterR")

library(tooltipsterR)
library(htmltools)
library(dplyr)
library(tidyr)
library(pipeR)

biggest <- formatter(
  "span",
  class = x ~ ifelse(rank(-x) ==1, "tooltipster-tooltip", ""),
  style = x ~ ifelse(rank(-x) ==1, "color:green;font-weight:bold;", NA),
  title = x ~ ifelse(rank(-x) ==1, "most common", NA)
)

hair_eye <- prop.table(
  xtabs(Freq~Eye+Hair,HairEyeColor)
) %>>%
  data.frame() %>>%
  mutate( Freq = percent(Freq) ) %>>%
  spread(
    Eye, Freq
  )

hair_eye %>>%
  (
    formattable(
      .,
      structure(
        lapply(
          colnames(.)[-1],
          function(z){
            biggest
          }
        ),
        names = colnames(.)[-1]
      )
    )
  ) %>>%
  formattable::as.htmlwidget() %>>%
  tagList(
    tooltipster()
  ) %>>%
  browsable

```

```{r "example02_run", echo = F}
#devtools::install_github("timelyportfolio/tooltipsterR")

library(tooltipsterR)
library(htmltools)
library(formattable)
library(dplyr)
library(tidyr)
library(pipeR)

biggest <- formatter(
  "span",
  class = x ~ ifelse(rank(-x) ==1, "tooltipster-tooltip", ""),
  style = x ~ ifelse(rank(-x) ==1, "color:green;font-weight:bold;", NA),
  title = x ~ ifelse(rank(-x) ==1, "most common", NA)
)

hair_eye <- prop.table(
  xtabs(Freq~Eye+Hair,HairEyeColor)
) %>>%
  data.frame() %>>%
  mutate( Freq = percent(Freq) ) %>>%
  spread(
    Eye, Freq
  )

hair_eye %>>%
  (
    formattable(
      .,
      structure(
        lapply(
          colnames(.)[-1],
          function(z){
            biggest
          }
        ),
        names = colnames(.)[-1]
      )
    )
  ) %>>%
  formattable::as.htmlwidget() %>>%
  tagList(
    tooltipster(".tooltipster-tooltip")
  ) %>>%
  save_tags(file="example02.html")
```

### With SVG

My favorite thing about `tooltipster` is it works with `SVG`.  Let's add a tooltip to our points that tells us their color.  By the way, if you have not seen [`svglite`](https://github.com/hadley/svglite), I strongly encourage checking [it](https://github.com/hadley/svglite) out.

```{r "example03", eval = F}
#devtools::install_github("timelyportfolio/tooltipsterR")

library(tooltipsterR)
library(htmltools)
#devtools::install_github("hadley/svglite")
library(svglite)

browsable(
  tagList(
    htmlSVG(plot(1:3,col=blues9[7:9],pch=16,cex=3)),
    tooltipster(),
    tags$script(
"
$('circle').each(function(){
  $(this).tooltipster({
    content: $(this).css('fill')
  })
})
"      
    )
  )
)
```

```{r "example03_run", echo = F}
#devtools::install_github("timelyportfolio/tooltipsterR")

library(tooltipsterR)
library(htmltools)
#devtools::install_github("hadley/svglite")
library(svglite)

tagList(
  htmlSVG(plot(1:3,col=blues9[7:9],pch=16)),
  tooltipster(),
  tags$script(
"
$('circle').each(function(){
$(this).tooltipster({
  content: $(this).css('fill')
})
})
"      
  )
) %>>%
  save_tags(file="example03.html")
```



## Thanks

Thanks [Caleb Jacob](http://calebjacob.com) for `tooltipster`.

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

