---
title: "Week03 - More Networks"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
---

```{r "setup",echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
  #, results = 'asis'
)

library(pipeR)
```

## htmlwidgets News This Week
@smartinsightsfromdata has started two new widgets over the past couple of weeks.

-[`rd3pie`](https://github.com/smartinsightsfromdata/rd3pie) - [`d3pie`](http://d3pie.org/) wrapper.  This should make interactive pie charts much easier than [my experiment](http://timelyportfolio.blogspot.com/2014/07/beer-and-pie-rcharts-pie-charts-with.html) from a couple months ago with `d3pie` + `rCharts`.

-[`rpivotTable`](https://github.com/smartinsightsfromdata/rpivotTable) - [`pivottable.js`](http://nicolas.kruchten.com/pivottable/examples/) wrapper.  `pivottable.js` from [\@nicolaskruchten](http://nicolas.kruchten.com/) is both nicely done and very active.  I had `pivottable.js` at the top of my list for a widget of the week.  I look forward to helping [\@smartinsightsfromdata](https://github.com/smartinsightsfromdata) push this into a full-fledged and documented `htmlwidget` assuming he accepts my offer.

@rich-iannone

[\@rich-iannone](https://github.com/rich-iannone) is unstoppable.  He continues to do amazing things with [`DiagrammeR`](https://github.com/rich-iannone/DiagrammeR).  Be sure to check out all the new documentaion, examples, and functionality.

## This Week's Widget - `hierNetwork` for [`networkD3`](https://github.com/christophergandrud/networkD3)

Unfortunately, this week's widget is still alpha.  `hierNetwork` is not a standalone widget.  Rather, it aims to add some new visualization types and interactivity to [`networkD3`](https://github.com/christophergandrud/networkD3).  Given the popularity of `htmlwidgets`, I think packaging/curating multiple widgets in a single R package will be very valuable.  These are my early thoughts on packaging `htmlwidgets`.


1. type of visualization
1. field of study or domain of application
1. similar Javascript dependencies
1. symbiotic features
1. type of parameters/inputs, such as diagram in DiagrammeR for both mermaid and graphviz
1. interest of the package author in support and ongoing maintenance


[`networkD3`](https://github.com/christophergandrud/networkD3) from [\@christophergandrud](https://github.com/christophergandrud) and [\@jjallaire](https://github.com/jjallaire) started back in [August](https://github.com/christophergandrud/networkD3/commit/6dc73683a7370c7640520ae7778799d6735bc2fa) just one month after the first [commit](https://github.com/christophergandrud/networkD3/commit/6dc73683a7370c7640520ae7778799d6735bc2fa) of `htmlwidgets`.  It is clearly one of the first ever `htmlwidgets`.

`hierNetwork` all started with this [issue](https://github.com/christophergandrud/networkD3/issues/14) requesting a phylogram implementation in `networkD3` from existing code.  I thought a phylogram would be fun, so I started down that path and got an early working [version](http://bl.ocks.org/timelyportfolio/raw/59acc3853b02e47e0dfc/).  However, a phylogram is really a specialized dendrogram, so I thought additional dendrogram-type functionality in `networkD3` might be more generally useful.

I still plan to work on phylograms, so if anyone is interested, please let me know.  For now, let's have a quick look at `networkD3` and the new `hierNetwork()`.

## What Does it Do?
The [visualizations from `networkD3`](http://christophergandrud.github.io/networkD3/) offer some really beautiful chart options for all sorts of network data.  It even has a `as.treeNetwork` function for easy conversion of `hclust` and `dendrogram` class to the very un-R-like nested data structure preferred by `d3`.  I'll prove how easy it is with two lines of code.


```{r "example01",eval = F}
library(networkD3)

# use this simple hclust example from
#  http://rpubs.com/gaston/dendrograms
hc = hclust(dist(mtcars))

treeNetwork( 
  as.treeNetwork(hc, "mtcars")
)
```
```{r "example01_run",echo = F}
library(networkD3)

hc = hclust(dist(mtcars))

treeNetwork( 
  as.treeNetwork(hc, "mtcars")
  , width = 750, height =750
) %>>%
      htmlwidgets::as.iframe(
        file="example01.html"
        , libdir = "lib"
        , selfcontained=F
        , width = 850
        , height =900
        , overflow = "hidden"
      )
```

It doesn't get much easier than that, but `d3` provides multiple [hierarchical layout](https://github.com/mbostock/d3/wiki/Hierarchy-Layout), and I'm sure each of you has an opinion on which is your favorite.  Also, a little pan/zoom and collapse interactivity makes us happy.  Since the hard part of data conversion is mostly already done by `networkD3` I went to work and started building this from scratch (well copy/paste from basically everywhere).  Fortunately, before I got too far along, I discovered this reusable `d3.chart` hierarchy layout [`d3.chart.layout`](https://github.com/bansaghi/d3.chart.layout).  It is dang near perfect for this and gives us a whole lot.

Let's take the same `hclust` that we named `hc` from above and look at each of the diagram types.

```{r "example02", eval = F}
library(htmltools) 

tagList(
  lapply(
    c("tree.cartesian"
      ,"tree.radial"
      ,"cluster.cartesian"
      ,"cluster.radial"
    )
    ,function(chartType){
      hierNetwork( as.treeNetwork(hc), type = chartType, zoomable = T, collapsible = T )
    }
  )
)
```

```{r "example02_run", echo = F}
library(htmltools) 

i = 1
tagList(
  lapply(
    c("tree.cartesian"
      ,"tree.radial"
      ,"cluster.cartesian"
      ,"cluster.radial"
    )
    ,function(chartType){
      hn <- hierNetwork( as.treeNetwork(hc), type = chartType, zoomable = T, collapsible = T, margin = list(left=100), height = 750, width = 750 ) %>>%
        htmlwidgets::as.iframe(
        file=paste0("example02_",i,".html")
        , libdir = "lib"
        , selfcontained=F
        , width = 850
        , height =900
        , overflow = "hidden"
      )
      i <<- i + 1
      return(list(
         tags$h4(chartType)
         ,hn
        )
      )
    }
  )
)
```

For those of you wondering if there is a difference between `cluster` and `tree`, collapse some of the nodes.  If you still don't see it, `cluster` extends to the outermost level ( much like `hang` in `R` when plotting dendrograms ) while `tree` does not.  Also, I hope you enjoyed the pan/zoom functionality, which is enabled by `zoomable=T` and `collapsible=T`.


```{r}
library(rpart)
library(partykit)
library(rCharts)
library(rlist)
library(pipeR)

#set up a little rpart as an example
rp <- rpart(
  hp ~ cyl + disp + mpg + drat + wt + qsec + vs + am + gear + carb,
  method = "anova",
  data = mtcars,
  control = rpart.control(minsplit = 4)
)

str(rp)

rpk <- as.party(rp)

## get meta information
rpk.text <- capture.output( print(rpk) ) %>>%
  ( .[grep( x = ., pattern = "(\\[)([0-9]*)(\\])")] ) %>>%
  strsplit( "[\\[\\|\\]]" , perl = T) %>>%
  list.map(
    tail(.,2) %>>%
      (
        data.frame(
          "id" = as.numeric(.[1])
          , description = .[2]
          , stringsAsFactors = F )
      )
  ) %>>% list.stack

# binding the node names from rpk with more of the relevant meta data from rp
# i don't think that partykit imports this automatically for the inner nodes, so i did it manually
rpk.text <- cbind(rpk.text, rp$frame)

# rounding the mean DV value
rpk.text$yval <- round(rpk.text$yval, 2)

# terminal nodes have descriptive stats in their names, so I stripped these out
# so the final plot wouldn't have duplicate data
rpk.text$description <- sapply(strsplit(rpk.text[,2], ":"), "[", 1)


dat = rapply(rpk$node,unclass,how="replace")
#fill in information at the root level for now
#that might be nice to provide to our interactive graph
dat$info = rapply(
  unclass(rpk$data)[-1]
  ,function(l){
    l = unclass(l)
    if( class(l) %in% c("terms","formula","call")) {
      l = paste0(as.character(l)[-1],collapse=as.character(l)[1])
    }          
    attributes(l) <- NULL
    return(l)
  }
  ,how="replace"
)
dat = jsonlite::toJSON(
  dat
  ,auto_unbox = T
)
# replace kids with children to ease d3
dat = gsub( x=dat, pattern = "kids", replacement="children")
# change id to node to ease d3; will replace with name later
dat = gsub ( x=dat, pattern = '"id":([0-9]*)', replacement = '"name":"node\\1","size":nodesize\\1' )
# calling the root node by the dataset name, but it might make more sense to call it
# "root" so that the code can be generalized
dat = sub (x = dat, pattern = "node1", replacement = "mtcars")
# replacing the node names from node1, node2, etc., with the extracted node names and metadata from
# rpk.text, and rp$table. 
for (i in 2:nrow(rpk.text)) {
  dat = sub (
    x = dat
    , pattern = paste("node", i, sep = "")
    , replacement = paste(
      rpk.text[i,2]
      , ", mean = ", rpk.text[i,7]
      , ", n = ", rpk.text[i,4]
      , sep = ""
    )
    , fixed = T
  )
  dat = sub (
    x = dat
    , pattern = paste("nodesize", i, sep = "")
    , replacement = rpk.text[i,4]
    , fixed = T
  )
}

# replace size of root or node1
dat = sub (
  x = dat
  , pattern = "nodesize1"
  , replacement = rpk.text[1,4]
  , fixed = T
)

hN <- hierNetwork( jsonlite::fromJSON(dat), collapsible = T )
# fromJSON does not translate well so manual override
hN$x$root = dat
lapply(
  c("pack.nested"
    ,"pack.flattened"
    ,"partition.arc"
    ,"partition.rectangle"
    ,"treemap"
  )
  ,function(chartType){
    hN$x$options$type = chartType
    return(hN) 
  }
)

```
