---
title: "Week16 - Games in R"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## htmlwidgets News This Week

Be **sure** to check out [`rcstatebin`](https://github.com/ramnathv/rcstatebin) from Ramnath Vaidyanathan, author of `htmlwidgets`.  For `rcstatebin`, Ramnath created both the underlying `d3js` library and `R` wrapper, so they integrate beautifully.  In creating the `d3js` side, he is able to contribute back to the JavaScript community from which we source so much.  Avid readers already know, but for all the newest and updated `htmlwidgets`, just do this [Github search](https://github.com/search?o=desc&q=htmlwidgets&s=indexed&type=Code), and you'll quickly get up to date.

## This Week's Widget - `gamer`

If you are like me, the last thing you need is another excuse to procrastinate/"research" while you are in `R`.  I thought making `gamer`--this week's `htmlwidget`--would help expand our minds to think beyond charts and tables when we think `htmlwidgets`.  Wouldn't you know it the prolific Yihui Xie from RStudio has already made the [`fun` package](http://yihui.name/en/2011/08/the-fun-package-use-r-for-fun/) if you want more games for `R`.

Earlier this month, I spotted [this post](http://cacodaemon.de/index.php?id=67) about [`EntangledClone`](https://github.com/Cacodaimon/EntangledClone), which is written in `TypeScript` with help from [`d3.js`](http://d3js.org).  It is nice in that it is fun, it is small, and it is responsive, and it is easily embeddable, so go ahead and put it in your next `rmarkdown` research paper.  Most of the work(not hard at all) wrapping `EntangledClone` was locally building the project.  Then, I just separated the instantiation piece and pasted it into `entangler.js`.


## Quick Installation

`gamer` is very unlikely to go to CRAN, so for now please install with `devtools::install_github`. 

```{r eval = F}
devtools::install_github("timelyportfolio/gamer")
```

## Example

I think one example is plenty.  Press `tab` for a new game.

```{r "example01", echo=T, eval = F}
library(gamer)

entangler( )
```

```{r "example01_run", echo=F, eval = T}
library(gamer)

entangler( height = 400, width = 400 ) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 520
      , height = 520
      , overflow = "hidden"
    )
  )
```

## Thanks

Thanks so much

- [`EntangledClone`](https://github.com/Cacodaimon/EntangledClone) from [Guido Krömer](http://www.cacodaemon.de/)
- [`d3.js`](http://d3js.org) from Mike Bostock
- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

