---
title: "Week01 - Easy Flowcharts and Diagrams"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
---

```{r "setup",echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
  #, results = 'asis'
)

library(DiagrammeR)
library(htmltools)
```

## Already Cheating?
It's just week 1 of the Building Widgets [widget-a-week challenge](http://www.buildingwidgets.com/blog/2015/1/2/can-i-commit), and some might think I am already cheating for two reasons:

1. This widget was started last week, and
1. I'm not the sole author.

However, these "transgressions" will likely occur throughout the year, since many widgets will be multi-week projects, and I very much like collaborating on widgets.

## This Week's Widget - [`DiagrammeR`](https://github.com/rich-iannone/DiagrammeR)
[`DiagrammeR`](https://github.com/rich-iannone/DiagrammeR) is authored by Air Quality Scientist [Richard Iannone](http://about.me/rich_i/) who has quite a record of [GitHub activity](https://github.com/rich-iannone#contributions-calendar).  If you are asking why I am involved, [`DiagrammeR`](https://github.com/rich-iannone/DiagrammeR) didn't start out as a [`htmlwidget`](http://htmlwidgets.org).  Since I along with > 5,000 other starrers love [`mermaid.js`](https://github.com/knsv/mermaid) and the `htmlwidgets` infrastructure is so compelling, I thought [`DiagrammeR`](https://github.com/rich-iannone/DiagrammeR) would be perfect as one of the first `htmlwidgets`.  For more details on the conversion, see [Widgets for Christmas](http://timelyportfolio.blogspot.com/2014/12/widgets-for-christmas.html) or for those more technically inclined inspect this [pull request](https://github.com/rich-iannone/DiagrammeR/pull/2).

### What Does it Do?
[`mermaid.js`](https://github.com/knsv/mermaid) offers easy diagrams through a markdown-like spec and leverages [`d3`](http://d3js.org) and [`dagre-d3`](https://github.com/cpettitt/dagre-d3).  The `mermaid.js` [site](http://knsv.github.io/mermaid/flowchart.html) offers quite a few examples of flowcharts, experimental sequence diagrams, and other diagrams.

### Examples in `R`
In general for all the widgets to come, I will try to create entirely new examples and put them in `R` contexts.  I will also try to convert examples from documentation of popular static `R` packages.

<h4>Example 1 - Stylish Nodes</h4>
```{r}
DiagrammeR("
  graph LR;
    A[rect]-- add style -->A2[rect + style];
    B{rhombus}---|+ some style|B2{rhombus + style};
    C(rounded);   D((circle));
    style A2 fill:#c12,stroke-width:5px;
    style B2 fill:none, stroke-dasharray:10;
")
```
