---
title: "Week33 - Contribution Week"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## This Week's Widget - Contributions

Since I make the rules for what counts as a `htmlwidget`, I'll create this new rule.

<blockquote>
A series of contributions to existing htmlwidgets counts as an htmlwidget in the weekly series.
</blockquote>

This week I made contributions to [`edgebundleR`](https://github.com/garthtarr/edgebundleR), [`sunburstR`](https://github.com/timelyportfolio/sunburstR), and [`leaflet`](https://github.com/rstudio/leaflet).  Thanks to the authors for all their work on these existing `htmlwidgets` and their willingness to accept my pull requests.

## edgebundleR

This StackOverflow [question](http://stackoverflow.com/questions/30708674/network-chord-diagram-woes-in-r/32260962#32260962) prompted me to try to add some new feaures to [Garth Tarr's](https://github.com/garthtarr) very nice [`edgebundleR`](https://github.com/garthtarr/edgebundleR).  I'll use my answer to the question as the example of the new functionality.

```{r "example01", eval = FALSE}
# devtools::install_github("garthtarr/edgebundleR")

library(edgebundleR)
library(igraph)
library(data.table)

d <- structure(list(ID = c("KP1009", "GP3040", "KP1757", "GP2243", 
                           "KP682", "KP1789", "KP1933", "KP1662", "KP1718", "GP3339", "GP4007", 
                           "GP3398", "GP6720", "KP808", "KP1154", "KP748", "GP4263", "GP1132", 
                           "GP5881", "GP6291", "KP1004", "KP1998", "GP4123", "GP5930", "KP1070", 
                           "KP905", "KP579", "KP1100", "KP587", "GP913", "GP4864", "KP1513", 
                           "GP5979", "KP730", "KP1412", "KP615", "KP1315", "KP993", "GP1521", 
                           "KP1034", "KP651", "GP2876", "GP4715", "GP5056", "GP555", "GP408", 
                           "GP4217", "GP641"),
                    Type = c("B", "A", "B", "A", "B", "B", "B", 
                             "B", "B", "A", "A", "A", "A", "B", "B", "B", "A", "A", "A", "A", 
                             "B", "B", "A", "A", "B", "B", "B", "B", "B", "A", "A", "B", "A", 
                             "B", "B", "B", "B", "B", "A", "B", "B", "A", "A", "A", "A", "A", 
                             "A", "A"),
                    Set = c(15L, 1L, 10L, 21L, 5L, 9L, 12L, 15L, 16L, 
                            19L, 22L, 3L, 12L, 22L, 15L, 25L, 10L, 25L, 12L, 3L, 10L, 8L, 
                            8L, 20L, 20L, 19L, 25L, 15L, 6L, 21L, 9L, 5L, 24L, 9L, 20L, 5L, 
                            2L, 2L, 11L, 9L, 16L, 10L, 21L, 4L, 1L, 8L, 5L, 11L), Loc = c(3L, 
                                                                                          2L, 3L, 1L, 3L, 3L, 3L, 1L, 2L, 1L, 3L, 1L, 1L, 2L, 2L, 1L, 3L, 
                                                                                          2L, 2L, 2L, 3L, 2L, 3L, 2L, 1L, 3L, 3L, 3L, 2L, 3L, 1L, 3L, 3L, 
                                                                                          1L, 3L, 2L, 3L, 1L, 1L, 1L, 2L, 3L, 3L, 3L, 2L, 2L, 3L, 3L)),
               .Names = c("ID", "Type", "Set", "Loc"), class = "data.frame",
               row.names = c(NA, -48L))

# let's add Loc to our ID
d$key <- d$ID
d$ID <- paste0(d$Loc,".",d$ID)

# Get vertex relationships
sets <- unique(d$Set[duplicated(d$Set)])
rel <-  vector("list", length(sets))
for (i in 1:length(sets)) {
  rel[[i]] <- as.data.frame(t(combn(subset(d, d$Set ==sets[i])$ID, 2)))
}

rel <- rbindlist(rel)

# Get the graph
g <- graph.data.frame(rel, directed=F, vertices=d)
clr <- as.factor(V(g)$Loc)
levels(clr) <- c("salmon", "wheat", "lightskyblue")
V(g)$color <- as.character(clr)
V(g)$size = degree(g)*5


edgebundle( g )
```

```{r "example01_run", eval = TRUE, echo = FALSE, fig.keep='none'}
# devtools::install_github("garthtarr/edgebundleR")

library(edgebundleR)
library(igraph)
library(data.table)

d <- structure(list(ID = c("KP1009", "GP3040", "KP1757", "GP2243", 
                           "KP682", "KP1789", "KP1933", "KP1662", "KP1718", "GP3339", "GP4007", 
                           "GP3398", "GP6720", "KP808", "KP1154", "KP748", "GP4263", "GP1132", 
                           "GP5881", "GP6291", "KP1004", "KP1998", "GP4123", "GP5930", "KP1070", 
                           "KP905", "KP579", "KP1100", "KP587", "GP913", "GP4864", "KP1513", 
                           "GP5979", "KP730", "KP1412", "KP615", "KP1315", "KP993", "GP1521", 
                           "KP1034", "KP651", "GP2876", "GP4715", "GP5056", "GP555", "GP408", 
                           "GP4217", "GP641"),
                    Type = c("B", "A", "B", "A", "B", "B", "B", 
                             "B", "B", "A", "A", "A", "A", "B", "B", "B", "A", "A", "A", "A", 
                             "B", "B", "A", "A", "B", "B", "B", "B", "B", "A", "A", "B", "A", 
                             "B", "B", "B", "B", "B", "A", "B", "B", "A", "A", "A", "A", "A", 
                             "A", "A"),
                    Set = c(15L, 1L, 10L, 21L, 5L, 9L, 12L, 15L, 16L, 
                            19L, 22L, 3L, 12L, 22L, 15L, 25L, 10L, 25L, 12L, 3L, 10L, 8L, 
                            8L, 20L, 20L, 19L, 25L, 15L, 6L, 21L, 9L, 5L, 24L, 9L, 20L, 5L, 
                            2L, 2L, 11L, 9L, 16L, 10L, 21L, 4L, 1L, 8L, 5L, 11L), Loc = c(3L, 
                                                                                          2L, 3L, 1L, 3L, 3L, 3L, 1L, 2L, 1L, 3L, 1L, 1L, 2L, 2L, 1L, 3L, 
                                                                                          2L, 2L, 2L, 3L, 2L, 3L, 2L, 1L, 3L, 3L, 3L, 2L, 3L, 1L, 3L, 3L, 
                                                                                          1L, 3L, 2L, 3L, 1L, 1L, 1L, 2L, 3L, 3L, 3L, 2L, 2L, 3L, 3L)),
               .Names = c("ID", "Type", "Set", "Loc"), class = "data.frame",
               row.names = c(NA, -48L))

# let's add Loc to our ID
d$key <- d$ID
d$ID <- paste0(d$Loc,".",d$ID)

# Get vertex relationships
sets <- unique(d$Set[duplicated(d$Set)])
rel <-  vector("list", length(sets))
for (i in 1:length(sets)) {
  rel[[i]] <- as.data.frame(t(combn(subset(d, d$Set ==sets[i])$ID, 2)))
}

rel <- rbindlist(rel)

# Get the graph
g <- graph.data.frame(rel, directed=F, vertices=d)
clr <- as.factor(V(g)$Loc)
levels(clr) <- c("salmon", "wheat", "lightskyblue")
V(g)$color <- as.character(clr)
V(g)$size = degree(g)*5


edgebundle( g, width = 500 ) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 620
      , height = 620
      , overflow = "hidden"
   ))
```


## sunburstR

[`sunburstR`](https://github.com/timelyportfolio/sunburstR) gets some `Shiny` support (see [issue](https://github.com/timelyportfolio/sunburstR/issues/3)) and various bug fixes thanks to help from [Keeghan Hines](https://github.com/keeganhines).  In addition, `sunburstR` received some new functionality (see [this issue](https://github.com/timelyportfolio/sunburstR/issues/2) for details).

## Leaflet.measure

I found this nifty little measuring control [Leaflet.measure](https://github.com/ljagis/leaflet-measure) for Leaflet, and I thought it would be a good reason for me to learn how to add plugins/controls to Leaflet.  See if you can measure Central Park.


```{r "example03", eval = FALSE}
# devtools::install_github("rstudio/leaflet")
library(leaflet)

leaf <- leaflet() %>%
  addTiles() %>%
  # central park
  fitBounds(-73.96, 40.75, -73.98,40.76) %>%
  addMeasure()

leaf
```

```{r "example03_run", eval = TRUE, echo = FALSE, fig.keep='none'}
 library(leaflet)

leaf <- leaflet(height = 500, width = 500) %>%
  addTiles() %>%
  # central park
  fitBounds(-73.92, 40.75, -73.97, 40.77) %>%
  addMeasure()

leaf%>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example03.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 620
      , height = 620
      , overflow = "hidden"
   ))
```


## Thanks

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

