---
title: "Week12 - Intense Image Viewer"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
  #, results = 'asis'
  , fig.width = 8
  , fig.height = 4
)


save_tags <- function (tags, file, selfcontained = F, libdir = "./lib") 
{
    if (is.null(libdir)) {
        libdir <- paste(tools::file_path_sans_ext(basename(file)), 
            "_files", sep = "")
    }
    htmltools::save_html(tags, file = file, libdir = libdir)
    if (selfcontained) {
        if (!htmlwidgets:::pandoc_available()) {
            stop("Saving a widget with selfcontained = TRUE requires pandoc. For details see:\n", 
                "https://github.com/rstudio/rmarkdown/blob/master/PANDOC.md")
        }
        htmlwidgets:::pandoc_self_contained_html(file, file)
        unlink(libdir, recursive = TRUE)
    }
    return(htmltools::tags$iframe(src= file, height = "400px", width = "100%", style="border:0;"))
}

library(pipeR)
```

## htmlwidgets News This Week

Just do this [Github search](https://github.com/search?o=desc&p=35&q=htmlwidgets&s=indexed&type=Code), and you'll likely see all the newest and latest.

## This Week's Widget - `imageR` | Image Viewing
While I consider [`svgPanZoom`](http://www.buildingwidgets.com/blog/2015/1/15/week-02-) as one of the most universally applicable `htmlwidgets`, sometimes non-svg images are just easier and more comfortable.  I spotted Tim Holman's [Intense Image Viewer](http://tholman.com/intense-images), and I thought it would be a nice start to an `htmlwidget` that eventually offers multiple viewing experiences for images and plots from R.  I very creatively called it `imageR`.

> **I'm bummed that I have to use iframes with this blog, so the expanded view is contained by the iframe size**.

To see the full effect, run `imageR` locally, or see the very nice examples from [Intense Image Viewer](http://tholman.com/intense-images).


### Quick Installation

`imageR` has not stabilized enough for CRAN, so for now please install with `devtools::install_github`.

```{r eval = F}
devtools::intall_github("timelyportfolio/imageR")
```


### Simple Example

I say this is simple, but getting the image right will require use of `base64` encoding, which might be new to you.  `base64` might sound hard, but fortunately, two packages `base64enc` (which we'll use) and `base64` make it easy.

```{r "example01", eval = FALSE}
library(imageR)
library(htmltools)
library(base64enc)

# make a tempfile to hold our chart  
tf <- tempfile()
# delete our tempfile on exit
on.exit(unlink(tf))

# save our chart as a big png; exaggerate the width
png( tf, height = 800, width = 1600 )
  # make a kind of pretty chart
  plot(sin(0:90),cos(seq(0,360,by=4)),type="l", col= topo.colors(4))
  lines(sin(0:90),cos(0:90),type="l",col=topo.colors(4)[3])
dev.off()

intense(
  tags$img(
    style="height:200px;"
    ,src = dataURI(file=tf,mime="image/png")
  )
  ,height = 200
)
```

```{r "example01_run", eval = TRUE, echo = FALSE}
library(imageR)
library(htmltools)
library(base64enc)

# ignore the local if it confuses you
# only use it to delete the tempfile on exit
local({
  # make a tempfile to hold our chart  
  tf <- tempfile()
  # delete our tempfile on exit
  on.exit(unlink(tf))
  
  # save our chart as a big png; exaggerate the width
  png( tf, height = 800, width = 1600 )
    # make a kind of pretty chart
    plot(sin(0:90),cos(seq(0,360,by=4)),type="l", col= topo.colors(4))
    lines(sin(0:90),cos(0:90),type="l",col=topo.colors(4)[3])
  dev.off()
  
  intense(
    tags$img(
      style="height:200px;"
      ,src = dataURI(file=tf,mime="image/png")
    )
    ,height = 200
  )  %>>% htmlwidgets::as.iframe(
      file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , height = 400
      , overflow = "hidden"
   )
})
```


### Example with Pairs to Show Auto

In the first example, we exaggerated the width so you could see how the image pans from right to left based on your mouse or touch movement.  `intense-images` automatically decides to pan.  To see how it adjusts resize your window when in full-screen display mode.

```{r "example02", eval = FALSE}
library(imageR)
library(htmltools)
library(base64enc)

# make a tempfile to hold our chart  
tf <- tempfile()
# delete our tempfile on exit
on.exit(unlink(tf))

# save our chart as a big png
png( tf, height = 1200, width = 1200 )
  # make a kind of pretty chart
  pairs(USJudgeRatings, main="Example from ?pairs")
dev.off()

intense(
  tags$img(
    style="height:400px;overflow:hidden;"
    ,src = dataURI(file=tf,mime="image/png")
  )
  ,height = 400
)
```


```{r "example02_run", eval = T, echo = F}
library(imageR)
library(htmltools)
library(base64enc)

# ignore the local if it confuses you
# only use it to delete the tempfile on exit
local({
  # make a tempfile to hold our chart  
  tf <- tempfile()
  # delete our tempfile on exit
  on.exit(unlink(tf))
  
  # save our chart as a big png
  png( tf, height = 1200, width = 1200 )
    # make a kind of pretty chart
    pairs(USJudgeRatings, main="Example from ?pairs")
  dev.off()
  
  intense(
    tags$img(
      style="height:400px;overflow:hidden;"
      ,src = dataURI(file=tf,mime="image/png")
    )
    ,height = 400
  )  %>>% htmlwidgets::as.iframe(
      file="example02.html"
      , libdir = "lib"
      , selfcontained=F
      , height = 450
      , overflow = "hidden"
   )
})
```

### Example with a Non-`img`

`intense-images` also works with non-`img` elements, such as a `div`.  To make it work correctly, we'll use some extra tag attributes `data-image`, `data-title`, and `data-caption` on our non-`img`.  I bet you can guess what these attributes mean.  We'll use the same plot with some different colors as above to illustrate the difference.

```{r "example03", eval = FALSE}
library(imageR)
library(htmltools)
library(base64enc)

# make a tempfile to hold our chart  
tf <- tempfile()
tf2 <- tempfile()
# delete our tempfile on exit
on.exit({unlink(tf);unlink(tf2)})

png( tf, height = 1200, width = 1200 )
  # make a kind of pretty chart
  pairs(USJudgeRatings, main="Example from ?pairs")
dev.off()

# save our chart as a smaller png than above for demo
png( tf2, height = 800, width = 800 )
  # make a kind of pretty chart
  pairs(USJudgeRatings, main="Example from ?pairs",
    pch = 4, col="blue", cex = 0.5
  )
dev.off()

tagList(
  intense(
    # method already shown
    tags$img(
      style="height:200px;width:200px;overflow:hidden"
      ,src = dataURI(file=tf,mime="image/png")
    )
    ,height=250, width = 400
  )
  , intense(
    tags$div(
      "data-image" = dataURI(file=tf2,mime="image/png")
      ,"data-title" = "?pairs example"
      ,"data-caption" = "neat things happen with extra attributes"
      ,style = "height:50px;width:250px;border: dashed 0.2em lightgray;"
      ,"not an img but still intense; click me"
    )
    ,height = 250, width = 50
  )
)
```

```{r "example03_run", eval = T, echo = F}
library(imageR)
library(htmltools)
library(base64enc)

# ignore the local if it confuses you
# only use it to delete the tempfile on exit
local({
  # make a tempfile to hold our chart  
  tf <- tempfile()
  tf2 <- tempfile()
  # delete our tempfile on exit
  on.exit({unlink(tf);unlink(tf2)})

  png( tf, height = 1200, width = 1200 )
    # make a kind of pretty chart
    pairs(USJudgeRatings, main="Example from ?pairs")
  dev.off()
  
  # save our chart as a smaller png than above for demo
  png( tf2, height = 800, width = 800 )
    # make a kind of pretty chart
    pairs(USJudgeRatings, main="Example from ?pairs",
      pch = 4, col="blue", cex = 0.5
    )
  dev.off()
  
  tagList(
    intense(
      # method already shown
      tags$img(
        style="height:200px;width:200px;overflow:hidden"
        ,src = dataURI(file=tf,mime="image/png")
      )
      ,height=250, width = 400
    )
    , intense(
      tags$div(
        "data-image" = dataURI(file=tf2,mime="image/png")
        ,"data-title" = "?pairs example"
        ,"data-caption" = "neat things happen with extra attributes"
        ,style = "height:50px;width:250px;border: dashed 0.2em lightgray;"
        ,"not an img but still intense; click me"
      )
      ,height = 250, width = 50
    )
  )  %>>% save_tags("example03.html")
})
```

### With Other htmlwidgets and Flickr

htmlwidgets are composeable.  Let's add `intense-images` to an example from [Week 10 | Responsive Toolbars](http://www.buildingwidgets.com/blog/2015/3/11/week-10-responsive-toolbars) with `navr` and `sortableR`.  You might say this is more like publishing with HTML, and you are right.  R is really not necessary for this example, but when you are in a context, it is often much more efficient to stay in that context.  Thanks

```{r "example04", eval = FALSE}
library(htmltools)
library(curl)
library(navr)
library(sortableR)
library(imageR)

n1 <- navr(
  selector = "#sortableR-toolbar"
  ,taglist = tagList(
    tags$ul(id = "sort-navr"
      ,style="line-height:120px; text-align:center; vertical-align:middle;"
      ,tags$li(
        style="border: solid 0.1em white;border-radius:100%;line-height:inherit;width:130px;height:130px;"
        , class="fa fa-binoculars fa-4x"
        #  attribution everywhere Creative Commons Flickr
        #  awesome picture by https://www.flickr.com/photos/12859033@N00/2288766662/
        , "data-image" = paste0(
          "data:image/jpg;base64,"
          ,base64enc::base64encode(
            curl("https://farm4.staticflickr.com/3133/2288766662_c40c168b76_o.jpg","rb")
          )
        )
        , "data-title" = "Binoculars, a working collection"
        , "data-caption" = "awesome picture courtesy Flickr Creative Commons
            <a href = 'https://www.flickr.com/photos/12859033@N00/2288766662/'>jlcwalker</a>"
      )        
      ,tags$li(
        style="border: solid 0.1em white;border-radius:100%;line-height:inherit;width:130px;height:130px;"
        , class="fa fa-camera fa-4x"
        #  attribution everywhere Creative Commons Flickr
        #  awesome picture by https://www.flickr.com/photos/s58y/5607717791
        , "data-image" = paste0(
          "data:image/jpg;base64,"
          ,base64enc::base64encode(
            curl("https://farm6.staticflickr.com/5309/5607717791_b030229247_o.jpg","rb")
          )
        )
        , "data-title" = "Leica IIIc converted to IIIf BD ST"
        , "data-caption" = "awesome picture courtesy Flickr Creative Commons
            <a href = 'https://www.flickr.com/photos/s58y/5607717791'>s58y</a>"
      )
    )
  )
)

tagList(
  tags$div(
    id = "sortableR-toolbar"
    ,style="width:300px;border: dashed 0.2em lightgray; float:left;"
    ,tags$h3("sortableR Icons for Intense Images")
    ,"These icons drag and drop. Click on them for an"
    ,tags$strong("intense")
    ,"result."
  )
  ,add_font_awesome(n1)
  ,sortableR("sort-navr")
  ,intense( selector = "#sort-navr li" )
)
```

```{r "example04_run", eval = T, echo = F}
library(htmltools)
library(curl)
library(navr)
library(sortableR)
library(imageR)

n1 <- navr(
  selector = "#sortableR-toolbar"
  ,taglist = tagList(
    tags$ul(id = "sort-navr"
      ,style="line-height:120px; text-align:center; vertical-align:middle;"
      ,tags$li(
        style="border: solid 0.1em white;border-radius:100%;line-height:inherit;width:130px;height:130px;"
        , class="fa fa-binoculars fa-4x"
        #  attribution everywhere Creative Commons Flickr
        #  awesome picture by https://www.flickr.com/photos/12859033@N00/2288766662/
        , "data-image" = paste0(
          "data:image/jpg;base64,"
          ,base64enc::base64encode(
            curl("https://farm4.staticflickr.com/3133/2288766662_c40c168b76_o.jpg","rb")
          )
        )
        , "data-title" = "Binoculars, a working collection"
        , "data-caption" = "awesome picture courtesy Flickr Creative Commons
            <a href = 'https://www.flickr.com/photos/12859033@N00/2288766662/'>jlcwalker</a>"
      )        
      ,tags$li(
        style="border: solid 0.1em white;border-radius:100%;line-height:inherit;width:130px;height:130px;"
        , class="fa fa-camera fa-4x"
        #  attribution everywhere Creative Commons Flickr
        #  awesome picture by https://www.flickr.com/photos/s58y/5607717791
        , "data-image" = paste0(
          "data:image/jpg;base64,"
          ,base64enc::base64encode(
            curl("https://farm6.staticflickr.com/5309/5607717791_b030229247_o.jpg","rb")
          )
        )
        , "data-title" = "Leica IIIc converted to IIIf BD ST"
        , "data-caption" = "awesome picture courtesy Flickr Creative Commons
            <a href = 'https://www.flickr.com/photos/s58y/5607717791'>s58y</a>"
      )
    )
  )
)

tagList(
  tags$div(
    id = "sortableR-toolbar"
    ,style="width:300px;border: dashed 0.2em lightgray; float:left;"
    ,tags$h3("sortableR Icons for Intense Images")
    ,"These icons drag and drop. Click on them for an"
    ,tags$strong("intense")
    ,"result."
  )
  ,add_font_awesome(n1)
  ,sortableR("sort-navr")
  ,intense( selector = "#sort-navr li" )
) %>>% save_tags("example_04.html")
```


## Thanks

Thanks so much for all the work by

- [Tim Holman](http://tholman.com) for intense-images and his other experiments
- Flickr users [s58y](https://www.flickr.com/photos/s58y) and [jlcwalker](https://www.flickr.com/photos/12859033@N00) for the photos used in the last example
- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R`

