---
title: "Week33 - RBioFabric"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## This Week's Widget - `RBioFabric`

[William Longabaugh (\@wjrl)](https://github.com/wjrl) developed a novel hairball-less way to visualize networks called [BioFabric](http://www.biofabric.org/) in his 2012 paper.

<blockquote>
Longabaugh, W.J.R. <a href = "http://www.biomedcentral.com/1471-2105/13/275">Combing the hairball with BioFabric: a new approach for visualization of large networks</a>. BMC Bioinformatics, 13:275, 2012.
</blockquote>

Since then, he has done an incredible amount to speed adoption by implementing BioFabric in Java, JavaScript (`d3.js`), and R.  His work, numerous examples, and discussion convinced me of the power of BioFabric, and I have been waiting for a week that I could commit enough time to attempt an implementation as an `htmlwidget`.  When I saw that the genius prolific node.js programmer [Max Ogden \@maxogden](https://github.com/maxogden/biofabric) had taken a first step toward extending and generalizing @wjrl's `d3.js` code, I decided to see if I could combine the `R` and JavaScript into an `htmlwidget`.

As I plunged through the code, I decided to approach this by using the much fuller featured `bioFabric` plotting function in `R` with all the power of `igraph` to do most of the data conversion and preparation in `R`.  I stripped this from the JavaScript, and instead use `d3.js` just to render the graph and layout.  By far, the [Java version](http://www.biofabric.org/#launch) is the fullest implementation of BioFabric with lots of helpful interactivity and analysis tools.  Using it for inspiration, I added crude pan/zoom and then some mouseover interaction in the JavaScript.  There is still a whole lot to do, and I would love help, ideas, comments, and suggestions from anyone willing or interested.

I want to make sure I at least spend another sentence commending and thanking William Longabaugh (@wjrl) for not only his brilliant idea but his incredible commitment to it.  I can only hope that this little `htmlwidget` will help bring a little bit more attention and help spread the word.

## Quick Installation

`RBioFabric` is not yet on CRAN, so for now please install with `devtools::install_github`.

```{r eval = F}
devtools::install_github("timelyportfolio/RBioFabric")
```

## Examples

Let's start simple with some `igraph` data.  It is beyond the scope of this post to explain BioFabric.  I highly encourage enjoying all of William's incredible [resources on BioFabric](http://biofabric.org) to get a better sense of how to interpret these special diagrams.

### Bridges of Koenigsberg

```{r "example01", eval = FALSE}
# devtools::install_github("timelyportfolio/RBioFabric")
library(igraph)
library(igraphdata)
library(RBioFabric)

data("Koenigsberg")

#  strange but we actually use the bioFabric plot function
#   to give us the data

bioFabric_htmlwidget( bioFabric( Koenigsberg) )
```

```{r "example01_run", eval = TRUE, echo = FALSE, fig.keep='none'}
library(igraph)
library(igraphdata)
library(RBioFabric)

data("Koenigsberg")

#  strange but we actually use the bioFabric plot function
#   to give us the data

bioFabric_htmlwidget( bioFabric( Koenigsberg ),width=530,height=300) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 700
      , height = 420
      , overflow = "hidden"
   ))
```

### shadowLinks

In the case of the Bridges of Koenigsberg, `shadowLinks` probably are not that helpful, but in bigger more complicated networks, `shadowLinks = TRUE` can help preserve and communicate a better sense of connectedness.

```{r "example02", eval = FALSE}
# devtools::install_github("timelyportfolio/RBioFabric")
library(igraph)
library(igraphdata)
library(RBioFabric)

data("Koenigsberg")

#  strange but we actually use the bioFabric plot function
#   to give us the data

bioFabric_htmlwidget(
  bioFabric( Koenigsberg, shadowLinks = TRUE )
)
```

```{r "example02_run", eval = TRUE, echo = FALSE, fig.keep='none'}
# devtools::install_github("timelyportfolio/RBioFabric")
library(igraph)
library(igraphdata)
library(RBioFabric)

data("Koenigsberg")

#  strange but we actually use the bioFabric plot function
#   to give us the data

bioFabric_htmlwidget(
  bioFabric( Koenigsberg, shadowLinks = TRUE )
  ,height = 400
  ,width = 550
) %>>%  (
   htmlwidgets::as.iframe(
      .
      ,file="example02.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 720
      , height = 520
      , overflow = "hidden"
   ))
```

### Ordering Nodes

By default, nodes are sorted by degree.  However, `bioFabric` allows custom sort order through a function with `orderFun` or as a vector with `userOrder`.  

```{r "example03", eval = FALSE}
# devtools::install_github("timelyportfolio/RBioFabric")
library(igraph)
library(igraphdata)
library(RBioFabric)
library(dplyr)

data(karate)
# ?fastgreedy.community

fc <- fastgreedy.community(karate)

# let's sort by community and then degree
bioFabric_htmlwidget(
  bioFabric(
    karate
    , userOrder = 
      data.frame(membership = membership(fc), degree = degree(karate)) %>%
        mutate( id = rownames(.) ) %>%
        arrange( membership, desc( degree ) ) %>%
        select( id ) %>%
        as.vector
  )
)

```

```{r "example03_run", eval = TRUE, echo = FALSE, fig.keep='none'}
# devtools::install_github("timelyportfolio/RBioFabric")
library(igraph)
library(igraphdata)
library(RBioFabric)
library(dplyr)

data(karate)
# ?fastgreedy.community

fc <- fastgreedy.community(karate)

# let's sort by community and then degree
bioFabric_htmlwidget(
  bioFabric(
    karate
    , userOrder = 
      data.frame(membership = membership(fc), degree = degree(karate)) %>%
        mutate( id = rownames(.) ) %>%
        arrange( membership, desc( degree ) ) %>%
        select( id ) %>%
        as.vector
  )
  ,height = 550
  ,width = 600
) %>>%  (
   htmlwidgets::as.iframe(
      .
      ,file="example03.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 720
      , height = 670
      , overflow = "hidden"
   ))
```

### Les Mis

No network visualization is complete unless it now includes Les Mis.  It's a little small in the blog, but this will give us a good opportunity to try out the pan and zoom.

```{r "example04", eval = FALSE}
# devtools::install_github("timelyportfolio/RBioFabric")
library(RBioFabric)
library(igraph)

# d3 example from BioFabric
#  source:   https://github.com/wjrl/D3BioFabric
#  example:  http://rawgit.com/wjrl/D3BioFabric/master/src/JustBioFabric.html

miserables <- jsonlite::fromJSON(
  "https://rawgit.com/wjrl/D3BioFabric/master/src/data/miserablesSimple.json"
)

mis_igraph <- graph.data.frame(
  d = miserables$links
  ,vertices = data.frame(
    id = as.character( 0:(nrow(miserables$nodes)-1) ) 
    ,name = miserables$nodes
    ,stringsAsFactors = F
  )
)

bioFabric_htmlwidget( bioFabric( mis_igraph ))
```

```{r "example04_run", eval = TRUE, echo = FALSE, fig.keep='none'}
 # devtools::install_github("timelyportfolio/RBioFabric")
library(RBioFabric)
library(igraph)

# d3 example from BioFabric
#  source:   https://github.com/wjrl/D3BioFabric
#  example:  http://rawgit.com/wjrl/D3BioFabric/master/src/JustBioFabric.html

miserables <- jsonlite::fromJSON(
  "https://rawgit.com/wjrl/D3BioFabric/master/src/data/miserablesSimple.json"
)

mis_igraph <- graph.data.frame(
  d = miserables$links
  ,vertices = data.frame(
    id = as.character( 0:(nrow(miserables$nodes)-1) ) 
    ,name = miserables$nodes
    ,stringsAsFactors = F
  )
)

bioFabric_htmlwidget( bioFabric( mis_igraph ), height = 550, width = 600) %>>%  (
   htmlwidgets::as.iframe(
      .
      ,file="example04.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 720
      , height = 670
      , overflow = "hidden"
   ))
```


## Thanks

Thanks so much to [William Longabaugh (\@wjrl)](https://github.com/wjrl) for a brilliant idea and his commitment to it.

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

