---
title: "Week37 - parsetR"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## This Week's Widget - parsetR

As far as I know, [`ggparallel`](https://cran.r-project.org/web/packages/ggparallel/) is the only way to make [Robert Kosara's parallel sets](http://kosara.net/papers/2006/Kosara_TVCG_2006.pdf) in `R`.  While `ggparallel` does a very nice job, this ridiculously amazing interactive `d3.js` [parallel sets implementation](https://www.jasondavies.com/parallel-sets/) by [Jason Davies](http://jasondavies.com) needs a `htmlwidget`.  When a JavaScript library is as well designed as [`d3-parsets`](https://github.com/jasondavies/d3-parsets), wrapping it as an `htmlwidget` can be very quick and straightforward.  In this case, I was able to achieve a working implementation in two hours documented on Twitter.  Of course for a `htmlwidget` to be really good and worthwhile to the `R` community, a lot of time still should be spent on documentation, examples, and testing to satisfy the ["What Makes a Good htmlwidget"](https://github.com/ramnathv/htmlwidgets/wiki/A-Good-htmlwidget) checklist.

## Installation

```{r eval = FALSE}
devtools::install_github("timelyportfolio/parsetR")
```

## Examples + Vignette

Making interactive `d3.js` parallel sets from `R` is now as simple as this.

```{r eval = FALSE}
library(parsetR)
parset(Titanic)
```

I'll iframe the vignette below for more examples.  If it doesn't look good/right, then here is the direct [link](http://timelyportfolio.github.io/buildingwidgets/week37/vignette.html).  I highly encourage **clicking, playing, and exploring the interactivity** in these examples.

<iframe src = "http://timelyportfolio.github.io/buildingwidgets/week37/vignette.html" seamless width = "80%" height = 600></iframe>

## Thanks

`parsetR` was easy due to all the effort by [Jason Davies](http://jasondavies.com) with [`d3-parsets`](https://www.jasondavies.com/parallel-sets/).  The world is a better place because of all Jason Davies other crazy awesome `d3.js` and JavaScript work.  Thanks!

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

