---
title: "Week29 - d3treeR 2"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## htmlwidgets News This Week

A **gallery is up** thanks to Ryan Hafen, but not "official".  As always, for all the newest and updated `htmlwidgets`, just do this [Github search](https://github.com/search?o=desc&q=htmlwidgets&s=indexed&type=Code).

## This Week's Widget - `d3treeR` Again

[d3treeR](https://github.com/timelyportfolio/d3treeR) was released last week in the post [Week 28 | d3treeR](http://www.buildingwidgets.com/blog/2015/7/17/week-28-d3treer).  We'll do `d3treeR` again but we'll make a version 2 `d3tree2` using the versions from Mike Bostock's [Zoomable Treemaps](http://bost.ocks.org/mike/treemap/) and Zan Armstrong's fork [Zoomable Treemaps with Color]](https://gist.github.com/zanarmstrong/76d263bd36f312cb0f9f).

This `d3tree2` version should work much better when there are multiple levels or nesting.

## Quick Installation

`d3treeR` is not yet on CRAN and won't be unless someone asks, so for now please install with `devtools::install_github`.  If you installed last week, you'll need to do this again.

```{r eval = F}
devtools::install_github("timelyportfolio/d3treeR")
```


## Examples


### `treemap` as data

As stated in the introduction, `d3treeR` is built to work seamlessly with `treemaps` from the `R` treemap package.

```{r "example01", eval = FALSE}
# devtools::install_github("timelyportfolio/d3treeR")
library(treemap)
library(d3treeR)

# first example from ?treemap

data(GNI2010)
d3tree(
  treemap(
     GNI2010,
     index=c("continent", "iso3"),
     vSize="population",
     vColor="GNI",
     type="value"
   )
  ,rootname = "World"
)

d3tree2(
  treemap(
     GNI2010,
     index=c("continent", "iso3"),
     vSize="population",
     vColor="GNI",
     type="value"
   )
  ,rootname = "World"
)
```

```{r "example01_run", eval = TRUE, echo = FALSE}
# devtools::install_github("timelyportfolio/d3treeR")
library(treemap)
library(d3treeR)

# first example from ?treemap

data(GNI2010)
d3tree(
  treemap(
     GNI2010,
     index=c("continent", "iso3"),
     vSize="population",
     vColor="GNI",
     type="value"
   )
  ,rootname = "World"
  ,height = 520
  ,width = 620
) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example01a.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 800
      , height = 670
      , overflow = "hidden"
   ))

d3tree2(
  treemap(
     GNI2010,
     index=c("continent", "iso3"),
     vSize="population",
     vColor="GNI",
     type="value"
   )
  ,rootname = "World"
  ,height = 520
  ,width = 620
) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example01b.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 800
      , height = 670
      , overflow = "hidden"
   ))
```

```{r "example02", eval = FALSE}
# devtools::install_github("timelyportfolio/d3treeR")
library(treemap)
library(d3treeR)

# last example from ?treemap

data(business)
business$employees.growth <- business$employees - business$employees.prev
d3tree(
  # Brewer's Red-White-Grey palette reversed with predefined range
  treemap(business,
          index=c("NACE1", "NACE2"),
          vSize="employees",
          vColor="employees.growth",
          type="value",
          palette="-RdGy",
          range=c(-30000,30000))
  ,rootname = "World"
)

d3tree2(
  # Brewer's Red-White-Grey palette reversed with predefined range
  treemap(business,
          index=c("NACE1", "NACE2"),
          vSize="employees",
          vColor="employees.growth",
          type="value",
          palette="-RdGy",
          range=c(-30000,30000))
  ,rootname = "World"
)
```

```{r "example02_run", eval = TRUE, echo = FALSE}
library(treemap)
library(d3treeR)

# last example from ?treemap

data(business)
business$employees.growth <- business$employees - business$employees.prev
d3tree(
  # Brewer's Red-White-Grey palette reversed with predefined range
  treemap(business,
          index=c("NACE1", "NACE2"),
          vSize="employees",
          vColor="employees.growth",
          type="value",
          palette="-RdGy",
          range=c(-30000,30000))
  ,rootname = "World"
  ,height = 520
  ,width = 620
) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example02a.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 800
      , height = 670
      , overflow = "hidden"
   ))

d3tree2(
  # Brewer's Red-White-Grey palette reversed with predefined range
  treemap(business,
          index=c("NACE1", "NACE2"),
          vSize="employees",
          vColor="employees.growth",
          type="value",
          palette="-RdGy",
          range=c(-30000,30000))
  ,rootname = "World"
  ,height = 520
  ,width = 620
) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example02b.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 800
      , height = 670
      , overflow = "hidden"
   ))
```

### `JSON` as data

We can use `JSON` or lists from `R` to provide our hierarchy.  Let's use the ubiquitous `flare.json`.

```{r "example03", eval = FALSE}
library(treemap)
library(d3treeR)

  # with flare.json from Mike Bostock

d3tree(
  "http://bl.ocks.org/mbostock/raw/4063582/raw/flare.json"
  ,id = "name"
  ,height = 520
  ,width = 620
)

d3tree2(
  "http://bl.ocks.org/mbostock/raw/4063582/raw/flare.json"
  ,id = "name"
  ,height = 520
  ,width = 620
)
```

```{r "example03_run", eval = TRUE, echo = FALSE}
library(treemap)
library(d3treeR)

  # with flare.json from Mike Bostock

d3tree(
  "http://bl.ocks.org/mbostock/raw/4063582/raw/flare.json"
  ,id = "name"
  ,height = 520
  ,width = 620
) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example03a.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 800
      , height = 670
      , overflow = "hidden"
   ))

d3tree2(
  "http://bl.ocks.org/mbostock/raw/4063582/raw/flare.json"
  ,id = "name"
  ,height = 520
  ,width = 620
) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example03b.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 800
      , height = 670
      , overflow = "hidden"
   ))
```

## Thanks

Thanks so much to Martijn Tennekes, Edwin de Jonge, Mike Bostock, and Zan Armstrong.

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

