---
title: "Week43 - emoji"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set( 
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

save_tags <- function (tags, file, selfcontained = F, libdir = "./lib") 
{
    if (is.null(libdir)) {
        libdir <- paste(tools::file_path_sans_ext(basename(file)), 
            "_files", sep = "")
    }
    htmltools::save_html(tags, file = file, libdir = libdir)
    if (selfcontained) {
        if (!htmlwidgets:::pandoc_available()) {
            stop("Saving a widget with selfcontained = TRUE requires pandoc. For details see:\n", 
                "https://github.com/rstudio/rmarkdown/blob/master/PANDOC.md")
        }
        htmlwidgets:::pandoc_self_contained_html(file, file)
        unlink(libdir, recursive = TRUE)
    }
    return(htmltools::tags$iframe(src= file, height = "400px", width = "100%", style="border:0;"))
}


library(pipeR)

```

## This Week's Widget - twemoji

For this week's `htmlwidget`, all I needed was an excuse, and that excuse came in the form of a [tweet](https://twitter.com/JennyBryan/status/656985008121384960) by [Jenny Bryan](http://www.stat.ubc.ca/~jenny/).  Of course, we need an emoji `htmlwidget`.  Since [`remoji`](https://github.com/richfitz/remoji) already brings some emoji power to R, I thought I would try to piggyback on it and reduce the future number of `R` packages by at least 1.  There are lots of very good JavaScript emoji libraries on Github.  I chose [`twemoji`](http://twitter.github.io/twemoji/) from Twitter for my first round of experiments.

**This needs help, so I'd love any help you can give in the form of feedback, ideas, feature requests, use cases, examples, ...**

## Installation

This is not on CRAN and not even in the `master` branch, so to install we will need some help from `devtools::install_github`.

```{r eval = FALSE}
devtools::install_github("timelyportfolio/remoji@twemoji")
```

## Examples

Examples this week are a little difficult, since emoji generally appear inline, and my blog setup doesn't really like inline `htmlwidgets`.  Here is a quick [example](http://bl.ocks.org/timelyportfolio/83488a9fe43f1ebe4b5c) from `rmarkdown`.

<iframe src = "http://bl.ocks.org/timelyportfolio/raw/83488a9fe43f1ebe4b5c/" seamless style= "height:600px; width:100%;"></iframe>


### Know the Unicode

If you really love emoji, you might already know the unicode value for it.

```{r "example01", eval = F}
#devtools::install_github("timelyportfolio/remoji@twemoji")

library(remoji)
library(stringi)
library(htmltools)

tagList(
  twemoji(svg = TRUE),
  tags$div(HTML("&#x1F4AA"),style="height:200px;width:200px")
)
```


```{r "example01_run", echo = F}
#devtools::install_github("timelyportfolio/remoji@twemoji")

library(remoji)
library(stringi)
library(htmltools)

tagList(
  twemoji(svg = TRUE),
  tags$div(HTML("&#x1F4AA"),style="height:200px;width:200px")
) %>>%
  save_tags(file="example01.html")
```

### Find with remoji

`remoji` offers the helpful `find_emoji` function so we don't have to memorize unicode.  I discovered the secret combination for use with `htmltools`.  It looks like this `HTML(stri_trans_general(emoji(flag),"any-hex/xml"))`.

```{r "example02", eval = F}
#devtools::install_github("timelyportfolio/remoji@twemoji")

library(remoji)
library(stringi)
library(htmltools)

tagList(
  lapply(
    find_emoji("flag"),
    function(flag){
      tags$div(
        tags$p(
          style = "font-size:150%",
          flag,
          tags$span(
            HTML(stri_trans_general(emoji(flag),"any-hex/xml"))
          )
        )
      )
    }
  )
)
```


```{r "example02_run", echo = F}
#devtools::install_github("timelyportfolio/remoji@twemoji")

library(remoji)
library(stringi)
library(htmltools)

tagList(
  lapply(
    find_emoji("flag"),
    function(flag){
      tags$div(
        tags$p(
          style = "font-size:150%",
          flag,
          tags$span(
            HTML(stri_trans_general(emoji(flag),"any-hex/xml"))
          )
        )
      )
    }
  )
) %>>%
  save_tags(file="example02.html")
```



## Thanks

Thanks Twitter for `twemoji`.

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

