---
title: "Week14 - Exporting Widgets"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

save_tags <- function (tags, file, selfcontained = F, libdir = "./lib", height="400px", width="100%") 
{
    if (is.null(libdir)) {
        libdir <- paste(tools::file_path_sans_ext(basename(file)), 
            "_files", sep = "")
    }
    htmltools::save_html(tags, file = file, libdir = libdir)
    if (selfcontained) {
        if (!htmlwidgets:::pandoc_available()) {
            stop("Saving a widget with selfcontained = TRUE requires pandoc. For details see:\n", 
                "https://github.com/rstudio/rmarkdown/blob/master/PANDOC.md")
        }
        htmlwidgets:::pandoc_self_contained_html(file, file)
        unlink(libdir, recursive = TRUE)
    }
    return(htmltools::tags$iframe(src= file, height = height, width = width, style="border:0;"))
}
```

## htmlwidgets News This Week

Just do this [Github search](https://github.com/search?o=desc&q=htmlwidgets&s=indexed&type=Code), and you'll likely see all the newest and latest.  As a follow-up to last week's widget `stmBrowser`, I thought this research on [applications of `stm` to MOOCs](http://andrewgelman.com/2015/04/08/can-teachers-large-online-classes-use-text-data-online-learners/) is especially interesting.

## This Week's Widget - `exportwidget`

It seems exporting widgets is a popular request.  This week's widget `exportwidget` offers an easy way to add an export to `PNG` button to other `htmlwidgets` or `SVG`.  It works, but it is more an experiment and conversation starter for the broader discussion at [htmlwidgets issue #95](https://github.com/ramnathv/htmlwidgets/issues/95).  There are lots of questions looking for answers.

- Why would you want to export an `htmlwidget`?
- How would you like to export an `htmlwidget` - clientside or headless?
- In what format would you like your exported `htmlwidget?

Thanks so much to the authors of [`canvg`](https://github.com/gabelerner/canvg), [`svg-crowbar`](https://github.com/NYTimes/svg-crowbar), and [`Download-File-JS`](https://github.com/PixelsCommander/Download-File-JS) on which `exportwidget` relies.


## Quick Installation

As with almost all widgets posted here, `exportwidget` is not on CRAN, so for now please install with `devtools::install_github`.  Given enough interest, I'm happy to put in the effort to make this or any others CRAN-worthy.  Just let me know.

```{r eval = F}
devtools::install_github("timelyportfolio/exportwidget")
```

## Examples

### Example With SVG + Custom Font

I'll start with what I think is the least likely use case of this widget but a demanding case that eliminates many of the `SVG` exporting JavaScript libraries.  To conduct the test and as an extra benefit and proof of functionality, I'll add a custom font from Google Fonts with `htmlDependency` from `htmltools`.  External fonts cause trouble in some of the other libraries I evaluated.

```{r "example01", echo=T, eval = F}
library(exportwidget)
library(htmltools)
library(pipeR)

tagList(
  '<svg id = "svg_to_export" width="400" height="200">
    <text x="50" y="100" text-anchor="start" dy="14" style="font-family:\'Indie Flower\';font-size:36pt;">
        Custom Fonts
    </text>
  </svg>' %>>%
    HTML
  ,export_widget( "svg" )
) %>>%
  attachDependencies(list(
    htmlDependency(
      name = "IndieFlower"
      ,version = "0.1"
      ,src = c(href='http://fonts.googleapis.com/css?family=Indie+Flower')
      ,stylesheet = ""
    )
  ))
```

```{r "example01_run", echo=F, eval = T}
library(exportwidget)
library(htmltools)
library(pipeR)

tagList(
  '<svg id = "svg_to_export" width="400" height="200">
    <text x="50" y="100" text-anchor="start" dy="14" style="font-family:\'Indie Flower\';font-size:36pt">
        Custom Fonts
    </text>
  </svg>' %>>%
    HTML
  ,export_widget( "svg" )
) %>>%
  attachDependencies(list(
    htmlDependency(
      name = "IndieFlower"
      ,version = "0.1"
      ,src = c(href='http://fonts.googleapis.com/css?family=Indie+Flower')
      ,stylesheet = ""
    )
  ))  %>>% save_tags("example01.html", height = "270px", width = "470px")
```

### Example with another widget `streamgraph`

[Alex Bresler](https://twitter.com/abresler/status/583050042099007489) assures me that streamgraphs are art, and thanks to [Bob Rudis \@hrbrmstr](http://rud.is/), there is an `htmlwidget` [`streamgraph`](http://rud.is/b/2015/03/12/streamgraph-htmlwidget-version-0-7-released-adds-support-for-markers-annotations/) for that.  Let's see how we can use `exportwidget` to show somebody our "art". The `ChickWeight` dataset is definitely not the best for `streamgraphs` but it is small and built-in, so we'll go with it despite these issues.

```{r "example2", echo = T, eval = F}
library(dplyr)
library(streamgraph)
library(htmltools)
library(exportwidget)

# use the built in ChickWeight data set; not ideal but built-in
ChickWeight %>%
  group_by( Diet, Time ) %>%
  summarise( MeanWeight = mean(weight) ) %>%
  streamgraph( "Diet", "MeanWeight", "Time", scale = "continuous" ) %>%
  sg_legend( show = TRUE ) %>%
  tagList(
    export_widget( )
  )
```
```{r "example2_run", echo = F, eval = T}
library(dplyr)
library(streamgraph)
library(htmltools)

# use the built in ChickWeight data set; not ideal but built-in
ChickWeight %>%
  group_by( Diet, Time ) %>%
  summarise( MeanWeight = mean(weight) ) %>%
  streamgraph( "Diet", "MeanWeight", "Time", scale = "continuous", height = 400, width = 600 ) %>%
  sg_legend( show = TRUE ) %>%
  tagList(
    export_widget( )
  ) %>>% save_tags("example02.html", height = "500px", width = "700px")
```

### Example with multiple widgets

Let's assume that we love our `htmlwidgets` and like to use lots of them on a page.  `export_widget` without arguments will try to apply its button to all the `htmlwidgets` it finds.  This could be bad.  For instance, I doubt we would want to export a `navr`.  If there are `htmlwidgets` in your page that you don't want to export, you'll probably want to be more specific in your `export_widget( )` and explicitly export just those `htmlwidgets`.  As an example, with `grViz` from `DiagrammeR`, you could `export_widget(".grViz")` or with the `streamgraph` from the previous example, `export_widget(".streamgraph")`.  Any `css` selector should work as an argument to `export_widget( )`.

Let's try it with [`DiagrammeR`](http://rich-iannone.github.io/DiagrammeR), [`rcdimple`](http://www.buildingwidgets.com/blog/2015/3/18/week-11-dimple-as-htmlwidget), and [`networkD3`](https://github.com/christophergandrud/networkD3).

```{r "example03", echo = T, eval = F}
library(htmltools)
library(DiagrammeR)
library(rcdimple)
library(networkD3)
library(exportwidget)

tagList(
  # A slightly more involved example
  #  using example from http://www.graphviz.org/pdf/dotguide.pdf
  #    "Drawing graphs with dot"
  #    Emden R. Gansner and Eleftherios Koutsofios and Stephen North
  #    January 5, 2015
  grViz('
  digraph G {
        size = "4,4";
        main [shape = box]; /* this is a comment */
        main -> parse [weight = 8];
        parse -> execute;
        main -> init [style = dotted];
        main -> cleanup;
        execute -> { make_string; printf}
        init -> make_string;
        edge [color = red]; // so is this
        main -> printf;
        node [shape = box, style = filled, color = ".7 .3 1.0"];
        execute -> compare;
  }
  ')
  ,dimple(
    mtcars
    , mpg ~ cyl
    , groups = "cyl"
    , type = "bubble"
  ) %>>%
    default_colors( streamgraph:::tableau_colors( "tableau10medium" ) ) %>>%
    set_bounds( x = "10%", y = "15%", width = "80%", height = "70%" )
  ,simpleNetwork(
    data.frame(
      Source = c("A", "A", "A", "A", "B", "B", "C", "C", "D")
      ,Target = c("B", "C", "D", "J", "E", "F", "G", "H", "I")
    )
    ,height = 400
    ,width = 400
  )
  ,export_widget()
)

```
```{r "example03_run", echo = F, eval = T}
library(htmltools)
library(DiagrammeR)
library(rcdimple)
library(networkD3)
library(exportwidget)

tagList(
  # A slightly more involved example
  #  using example from http://www.graphviz.org/pdf/dotguide.pdf
  #    "Drawing graphs with dot"
  #    Emden R. Gansner and Eleftherios Koutsofios and Stephen North
  #    January 5, 2015
  grViz('
  digraph G {
        size = "4,4";
        main [shape = box]; /* this is a comment */
        main -> parse [weight = 8];
        parse -> execute;
        main -> init [style = dotted];
        main -> cleanup;
        execute -> { make_string; printf}
        init -> make_string;
        edge [color = red]; // so is this
        main -> printf;
        node [shape = box, style = filled, color = ".7 .3 1.0"];
        execute -> compare;
  }
  ', height = 400, width = 400)
  ,dimple(
    mtcars
    , mpg ~ cyl
    , groups = "cyl"
    , type = "bubble"
    , height = 400, width = 400
  ) %>>%
    default_colors( streamgraph:::tableau_colors( "tableau10medium" ) ) %>>%
    set_bounds( x = "10%", y = "15%", width = "80%", height = "70%" )
  ,simpleNetwork(
    data.frame(
      Source = c("A", "A", "A", "A", "B", "B", "C", "C", "D")
      ,Target = c("B", "C", "D", "J", "E", "F", "G", "H", "I")
    )
    ,height = 400
    ,width = 400
  )
  ,export_widget()
) %>>%
  save_tags("example03.html", height = "900px", width = "600px")
```

### Example | Headless with `webshot`

Often, you might not want `exportwidget` at all.  We very well might like a headless way to make a static copy of an `htmlwidget`.  For this, we can use the excellent package [`webshot`](https://github.com/wch/webshot) by Winston Chang from RStudio.  `webshot` uses the very widely used and tested `phantomjs`.  We can push an `htmlwidget` through the pipe and get a `PNG` out the other end.  Note, `webshot` and Windows aren't best friends yet.  If you suffer through Windows like I do, then you can install my forked `webshot`.

```{r "example04", echo = T, eval = F}
library(networkD3)
library(webshot)

html_print(
  simpleNetwork(
    data.frame(
      Source = c("A", "A", "A", "A", "B", "B", "C", "C", "D")
      ,Target = c("B", "C", "D", "J", "E", "F", "G", "H", "I")
    )
    ,height = 400
    ,width = 400
)) %>>%
  normalizePath(.,winslash="/") %>%
  gsub(x=.,pattern = ":/",replacement="://") %>%
  paste0("file:///",.) %>%
  webshot( file = "headless_widget.png", delay = 2 )
```

```{r "example04_run", echo = F, eval = T}
library(webshot)

html_print(
  simpleNetwork(
    data.frame(
      Source = c("A", "A", "A", "A", "B", "B", "C", "C", "D")
      ,Target = c("B", "C", "D", "J", "E", "F", "G", "H", "I")
    )
    ,height = 400
    ,width = 400
)) %>>%
  normalizePath(.,winslash="/") %>%
  gsub(x=.,pattern = ":/",replacement="://") %>%
  paste0("file:///",.) %>%
  webshot( file = "headless_widget.png", delay = 2, selector = ".html-widget-static-bound" )

tags$img(src = base64enc::dataURI(file="./headless_widget.png",mime="image/png"))
```

## Thanks

Thanks so much for all the work by

- [`canvg`](https://github.com/gabelerner/canvg) from Gabe Lerner
- [`svg-crowbar`](https://github.com/NYTimes/svg-crowbar) from the NY Times
- [`Download-File-JS`](https://github.com/PixelsCommander/Download-File-JS) from Denis Radin
- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R`

