---
title: "Week35 - gifrecordeR"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## This Week's Widget - gifrecordeR

Animated GIFs are really useful for documenting, illustrating, sharing,  and sometimes annoying.  [Yaron Noveh's](https://twitter.com/YaronNaveh) absolutely amazing [gifw00t]( https://github.com/yaronn/gifw00t/) lets us record GIFs in the browser client-side, so if I wrap it in a `htmlwidget` I can record GIFs of my RStudio Viewer or Shiny session and share them with the world.

**SVG support is currently weak, but a work-in-progress.**

## Installation

```{r eval = FALSE}
devtools::install_github("timelyportfolio/gifrecordeR")
```

## Examples

Wrestling with the SVG support and trying to squeeze these into a `htmlwidget` provde far more difficult than I imagined, so I ran out of time to write up a full post.  I hope you don't mind, but I just `iframed` the Readme.md for now.  I'll try to edit and make a full post for now, but the weekend holiday is starting and the family is up.   Also, I want to remind you of the weakly received Week 14's [exportwidget](http://www.buildingwidgets.com/blog/2015/4/9/week-14-exporting-widget) as a non-animated export option.

<iframe src = "http://timelyportfolio.github.io/gifrecordeR/README.html" height = 600 width = 700></iframe>

## Thanks

I'm still amazed we can do this.  Thanks so much to [Yaron Naveh \@yaronn](https://twitter.com/YaronNaveh) for making `gifw00t`.  Be sure to also check out [blessed-contrib](https://github.com/yaronn/blessed-contrib).

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

