---
title: "Week21 - Radial Tree"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

as.iframe <- function(widget, file = NULL, selfcontained = TRUE,
   libdir = NULL, width = '100%', height = 500, `border-width` = '0px', ...){
  if (is.null(file)){
    file <- basename(tempfile(fileext = '.html'))
  }
  htmlwidgets::saveWidget(widget,
                          file = file, selfcontained = selfcontained, libdir = libdir
  )
  styles = list(`border-width` = `border-width`, ...)
  style = paste(
    paste(paste(names(styles), styles, sep = ':'), collapse = ';'),
    ";"
  )
  content <- if (selfcontained){
    on.exit(unlink(file), add = TRUE)
    list(
      srcdoc = paste(readLines(file), collapse = '\n'),
      width = width, height = height, style = style
    )
  } else {
    list(
      src = file,
      width = width, height = height, style = style
    )
  }
  htmltools::tag('iframe', content)
}

```

## htmlwidgets News This Week

Avid readers already know, but for all the newest and updated `htmlwidgets`, just do this [Github search](https://github.com/search?o=desc&q=htmlwidgets&s=indexed&type=Code), and you'll quickly get up to date.

## This Week's Widget - `radialNetwork`

When I saw Wm Leler demonstrate his radial tree on this [d3.bayarea() Show and Tell](https://youtu.be/ewJdCnO9eQo?t=2m13s), I instantly wanted it as an `htmlwidget`.  His statement that with all the `d3` examples out there very few are full-featured explains perfectly why his [Radial Tree with UI](http://bl.ocks.org/wmleler/a734fb2bb3319a2cb386) is exceptional.  Unfortunately, I had to remove a couple of the features temporarily due to my inability to get them to work within the constraints of a widget.  I'll try to add them back as soon as I can.  Even with the removals, I think this radial tree will be a very useful addition to `networkD3`.

## Quick Installation

`radialNetwork` started as a fork of `networkD3`, but is now its own package.  To try it out, please use `devtools::install_github`.

```{r eval = F}
devtools::install_github("timelyportfolio/radialNetworkR")
```

## Example

While we are exploring `radialNetwork`, I would also like to publicize the helpful [`data.tree`](https://github.com/gluc/data.tree) package from Christopher Glur.  He has recently added the function to easily go from `data.tree` to the `JSON` network format expected by `d3.js`.  Let's see `data.tree` working with `radialNetwork`.

```{r "example01", echo=T, eval = F}
# you'll need this fork of networkD3
# devtools::install_github("timelyportfolio/radialNetworkR")
library(radialNetworkR)

# devtools::install_github( "gluc/data.tree" )
library(data.tree)

# use the acme data from data.tree
data(acme)

# unname = TRUE to get in the proper d3 format
radialNetwork( ToListExplicit(acme, unname = TRUE ) )

```

```{r "example01_run", echo=F, eval = T}
library(radialNetworkR)

# devtools::install_github( "gluc/data.tree" )
library(data.tree)

# use the acme data from data.tree
data(acme)

# unname = TRUE to get in the proper d3 format
radialNetwork(
  ToListExplicit(acme, unname = TRUE ),
  height= 500, width = 650
) %>>%
  (
    as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 800
      , height = 700
      , overflow = "hidden"
    )
  )
```

Of course, no `d3.js` network example is complete without `flare.json`.

```{r "example02", echo=T, eval = F}
# you'll need this fork of networkD3
# devtools::install_github("timelyportfolio/networkD3@feature/radialNetwork")
library(radialNetworkR)

radialNetwork(
  jsonlite::fromJSON(
   paste0(
     readLines(
       paste0(
         "https://gist.githubusercontent.com/mbostock/"
         ,"1093025/raw/a05a94858375bd0ae023f6950a2b13fac5127637/flare.json"
       )
     )
     ,collapse = "\n"
   ),
   simplifyDataFrame = FALSE
  )
)
```

```{r "example02_run", echo=F, eval = T}
# you'll need this fork of networkD3
# devtools::install_github("timelyportfolio/networkD3@feature/radialNetwork")
library(radialNetworkR)

radialNetwork(
  jsonlite::fromJSON(
   paste0(
     readLines(
       paste0(
         "https://gist.githubusercontent.com/mbostock/"
         ,"1093025/raw/a05a94858375bd0ae023f6950a2b13fac5127637/flare.json"
       )
     )
     ,collapse = "\n"
   ),
   simplifyDataFrame = FALSE
  )
, height= 600, width = 680 ) %>>%
  (
    as.iframe(
      .
      ,file="example02.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 800
      , height = 800
      , overflow = "hidden"
    )
  )
```

## Thanks

Thanks so much

- Wm Leler for his [Radial Tree with UI](http://bl.ocks.org/wmleler/a734fb2bb3319a2cb386) on which this is based
- [`d3.js`](http://d3js.org) from Mike Bostock
- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

