---
title: "Week13 - Interactive stm"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)
```

## htmlwidgets News This Week

Just do this [Github search](https://github.com/search?o=desc&q=htmlwidgets&s=indexed&type=Code), and you'll likely see all the newest and latest.

## This Week's Widget - `stmBrowser` | htmlwidgetized stmBrowser 

I get sad when I see a really great R package that generates `HTML/JS` that does not use `htmlwidgets`, but it is entirely understandable since `htmlwidgets` is so new.  This happened a couple days ago when I spotted a tweet announcing [`stmBrowser`](https://github.com/mroberts/stmBrowser) which pairs with [`stm`](http://structuraltopicmodel.com/) and [`stmCorrViz`](https://github.com/AntonioCoppola/stmCorrViz).  These related projects had already aroused my interest when I saw them appear on CRAN.  I felt like I should intervene early, so I asked for permission to convert, and the authors very generously responded with a "Yes. How can we help?"  Heartened by the quick positive response and the very nice `d3.js` code, I started banging away, and I think I'm close to an acceptable first iteration.

**Please note that any screwups or malfunctions are mine and not the original authors.**

### Quick Installation

The `htmlwidget` version of `stmBrowser` currently exists as a fork with a different branch on Github.  For now please install with `devtools::install_github`.  Also, you'll need [`stm`](http://cran.r-project.org/web/packages/stm/index.html) from CRAN.

```{r eval = F}
install.packages("stm")
devtools::install_github("timelyportfolio/stmBrowser@htmlwidget")
```


### Example in a Presentation

If I don't talk about Structural Topic Models (stm), maybe no one will discover that I don't have a clue.  Because of my ignorance, I'll use one of the examples from `?stm::stm`.  I'll slap it in a `rmarkdown` presentation to prove that it really is an `htmlwidget`.  This will also demonstrate one of the many benefits of `htmlwidgets`.

<iframe src = "http://timelyportfolio.github.io/buildingwidgets/week13/stm_presentation.html" style = "height: 600px; width:100%; border: " seamless></iframe>


## Thanks

Thanks so much for all the work by

- [Michael Freeman](http://mfviz.com), [Jason Chuang](http://jason.chuang.info), [Molly Roberts](http://margaretroberts.net), [Brandon Stewart](http://scholar.harvard.edu/bstewart) and [Dustin Tingley](http://scholar.harvard.edu/dtingley) for letting me barge in on their great work and mess it up in who knows how many ways
- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R`

