---
title: "Week02 - Pan and Zoom Almost Any R Graph"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
---

```{r "setup",echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
  #, results = 'asis'
)

library(pipeR)
```

## htmlwidgets News This Week
@yutannihiliation has been prolific this week with a couple new widget seeds and contributions to this widget [`metricsgraphics`](http://hrbrmstr.github.io/metricsgraphics/) announced last week.

- [`chartist`](https://github.com/yutannihilation/chartist) - [Chartist.js](http://gionkunz.github.io/chartist-js) wrapper
- [`gneisschart`](https://github.com/yutannihilation/gneisschart) - Gneisschart/[Chartbuilder (Quartz)](http://quartz.github.io/Chartbuilder/) wrapper

@dcenergy

- [`rflot`](https://github.com/dcenergy/rflot) - [Flot](https://github.com/dcenergy/rflot) wrapper

@armish

- [`coffeewheel`](https://github.com/armish/coffeewheel) - [`coffeewheel` by Jason Davies](https://www.jasondavies.com/coffee-wheel/) wrapper
- [`metabologram`](https://github.com/armish/metabologram) - a specialized form of `coffeewheel` mentioned above

Not launched last week, but a big miss by me is this `htmlwidget` for mapping from @lmullen.  It was one of the first `htmlwidgets` built after the Dec. 17 release announcement, and offers some nice functionality and interesting examples.

- [`cartographer`](https://github.com/lmullen/cartographer) - [`d3.carto.map`](https://github.com/emeeks/d3-carto-map) wrapper by @lmullen

Also, be sure to check out the new `graphviz` functionality added to last week's widget of the week [`DiagrammeR`](https://github.com/rich-iannone/DiagrammeR).

## This Week's Widget - [`svgPanZoom`](https://github.com/timelyportfolio/svgPanZoom)
I wanted to head in a slightly different direction with this week's widget.  Instead of wrapping Javascript charting libraries as most `htmlwidgets` to date have done, I wanted to let `R` still do the plotting, since so many of us know and love base `graphics`, `ggplot2`, and `lattice`.  I hope nearly every `R` user can easily incorporate this little widget into their `R` graphical analysis and publishing workflow.

## What Does it Do?
With lots of help from [`SVGAnnotation`](http://www.omegahat.org/SVGAnnotation/) or [`gridSVG`](https://sjp.co.nz/projects/gridsvg/), `svgPanZoom` will add pan and zoom functionality to nearly all `R` graphics using the [`svg-pan-zoom.js`](https://github.com/ariutta/svg-pan-zoom) tiny and dependency-free Javascript library.  Maybe a little graphic drawn from last week's `DiagrammeR` can help explain the process.

```{r}
library(DiagrammeR)

DiagrammeR("
  graph LR;
    gr[<code>graphics</code>]-->sp[<code>svgPlot</code>]
    la[<code>lattice/trellis</code>]-->sp
    la[<code>lattice/trellis</code>]-->gs[<code>gridSVG</code>]
    gg[<code>ggplot2</code>]-->sp
    gg[<code>ggplot2</code>]-->gs
    sp-->|+svg-pan-zoom.js|ht[Interactive Graphic]
    gs-->|+svg-pan-zoom.js|ht
    ht-->Shiny
    ht-->rs[RStudio Viewer]
    ht-->rm[R Markdown]
    ht-->html[standalone HTML]
    ht-->iframe
", height = 300, width = 500) %>>%
      htmlwidgets:::as.iframe(
        file="diagram.html"
        , libdir = "lib"
        , selfcontained=F
        , width = 600
        , height =400
        , overflow = "hidden"
      )