---
title: "Week37 - parsetR"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## This Week's Widget - parsetR


## Installation

```{r eval = FALSE}
devtools::install_github("timelyportfolio/parsetR")
```

## Examples + Vignette

Making interactive `d3.js` parallel sets is now as simple as this.

```{r eval = FALSE}
library(parsetR)
parset(Titanic)
```

I'll iframe the vignette below for more examples.  If it doesn't look good/right, then here is the direct [link](http://timelyportfolio.github.io/buildingwidgets/week37/vignette.html).

<iframe src = "http://timelyportfolio.github.io/buildingwidgets/week37/vignette.html" seamless width = "80%" height = 600></iframe>

## Thanks

`parsetR` was easy due to all the effort by [Jason Davies](http://jasondavies.com) with [`d3-parsets`](https://www.jasondavies.com/parallel-sets/).  The world is a better place because of all Jason Davies other crazy awesome `d3.js` and JavaScript work.  Thanks!

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

