---
title: "Week50 - functionplotR"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set( 
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```


## This Week's Widget - functionplotR

[Mauricio Poppe](https://twitter.com/iMauricio) has built a really amazing 2d `d3-js`-based function plotter [function-plot](https://github.com/maurizzzio/function-plot).  I think it has all sorts of helpful uses in `R`, so [functionplotR](https://github.com/timelyportfolio/functionplotR) will be this week's `htmlwidget`.  Note, I released a couple days early.  I hope nobody minds.

## Installation

This is not on CRAN, so to install we will need some help from `devtools::install_github`.

```{r eval = FALSE}
devtools::install_github("timelyportfolio/functionplotR")
```

## Examples

I was able to replicate nearly all of the [fantastic examples](http://maurizzzio.github.io/function-plot/) from the `function-plot` JavaScript library.  For the sake of brevity, I'll only include a couple here.  Go [here](https://github.com/timelyportfolio/functionplotR/blob/master/inst/examples/examples.R) or `?functionplot` to see all of them.


### One Simple Equation

```{r "example01", eval = F}
# devtools::install_github("timelyportfolio/functionplotR")

library(functionplotR)

functionplot("x + 2")
```

```{r "example01_run", echo = F}
# devtools::install_github("timelyportfolio/functionplotR")

library(functionplotR)

functionplot("x + 2", height = 300, width = 400 ) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 570
      , height = 470
      , overflow = "hidden"
    )
  )
```

### Multiple Simple Functions

There a couple ways to supply functions to `functionplot`.  Below is the easiest but least flexible.

```{r "example02", eval = F}
# devtools::install_github("timelyportfolio/functionplotR")

library(functionplotR)

functionplot(c("x + 2","x"))
```

```{r "example02_run", echo = F}
# devtools::install_github("timelyportfolio/functionplotR")

library(functionplotR)

functionplot(c("x + 2","x"), height = 300, width = 400 ) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example02.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 570
      , height = 470
      , overflow = "hidden"
    )
  )
```

### With Some Options

`function-plot` gives us lots of `options`.  Let's use some of them to configure our plot from above.


```{r "example03", eval = F}
# devtools::install_github("timelyportfolio/functionplotR")

library(functionplotR)

functionplot(
  # use the more flexible but a little more confusing
  #  way to provide functions
  fn = list(
    list( fn = "x + 2", color = "black" ),
    list( fn = "x", color = "gray" )
  ),
  yDomain = c(-4, 4),
  xDomain = c(-4, 4),
  width = 420, 
  height = 400
)
```

```{r "example03_run", echo = F}
# devtools::install_github("timelyportfolio/functionplotR")

library(functionplotR)

functionplot(
  # use the more flexible but a little more confusing
  #  way to provide functions
  fn = list(
    list( fn = "x + 2", color = "black" ),
    list(
      fn = "x",
      color = "gray",
      nSamples = 100,
      graphType = "scatter"
    )
  ),
  yDomain = c(-4, 4),
  xDomain = c(-4, 4),
  width = 420, 
  height = 400
) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example03.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 590
      , height = 570
      , overflow = "hidden"
    )
  )
```

### Function with Derivative

One of my favorite features of `function-plot` is the ability to plot interactive derivatives.  It does require you to know the derivative though.  With a quick Google, I think I now remember the derivative of `cos(x)`.  It is fun to put in something other than the derivative.  Try `x` as an example.

```{r "example04", eval = F}
# devtools::install_github("timelyportfolio/functionplotR")

library(functionplotR)

functionplot(
  fn = list(
    list(
      fn = "cos(x)",
      derivative = list(
        fn = "-sin(x)",
        updateOnMouseMove = TRUE
      )
    )
  )
)
```

```{r "example04_run", echo = F}
# devtools::install_github("timelyportfolio/functionplotR")

library(functionplotR)

functionplot(
  fn = list(
    list(
      fn = "cos(x)",
      derivative = list(
        fn = "-sin(x)",
        updateOnMouseMove = TRUE
      )
    )
  ),
  width = 420, 
  height = 400
) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example04.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 590
      , height = 570
      , overflow = "hidden"
    )
  )
```


## Thanks

Thanks [Mauricio Poppe](http://maurizzzio.com/) for the really special [`function-plot`](https://github.com/maurizzzio/function-plot/) plotting library.

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

