---
title: "Week50 - summarytrees"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set( 
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## Bleeding Edge - crosstalk

**I feel obligated to mention this again**.  If you want to get a preview of the exciting changes for `htmlwidgets`, be sure to check out `htmlwidgets` [issue 86](https://github.com/ramnathv/htmlwidgets/issues/86#issuecomment-159863989).  Thanks to Joe Cheng from RStudio for taking the lead on this.

We can see some early [examples](https://mytinyshinys.shinyapps.io/premierLeague) with `plotly` from early adopter [Andrew Clark \@pssguy](https://twitter.com/pssguy).

## This Week's Widget - summarytrees_htmlwidget

[Kenneth Shirley](http://www.kennyshirley.com/) and Howard Karloff authored an interesting [paper](https://scholar.google.com/scholar?cluster=11439479883993290700&hl=en&as_sdt=0,1).

<blockquote>
(2013). Howard Karloff and Kenneth E. Shirley. "Maximum Entropy Summary Trees", Computer Graphics Forum (Proc. EuroVis), Volume 32, Issue 3, Part 1, pp. 71-80.
</blockquote>

Subsequently, Kenny made an `R` package [`summarytrees`](https://github.com/kshirley/summarytrees) to create and visualize these summary trees.  I learned about the R/d3.js combination after Kenny's [presentation at JSM 2015](https://www.amstat.org/meetings/jsm/2015/onlineprogram/AbstractDetails.cfm?abstractid=314588).  Similar to the `htmlwidget` conversions of [`stmBrowser`](http://www.buildingwidgets.com/blog/2015/4/2/week-13-interactive-stm) and [`stmCorrViz`](http://www.buildingwidgets.com/blog/2015/12/1/week-48-stmcorrviz), I thought `summarytrees` would also be a great candidate for an `htmlwidget`.

## Installation

This is not on CRAN and only exists in a `summarytrees` fork, so to install we will need some help from `devtools::install_github`.

```{r eval = FALSE}
devtools::install_github("timelyportfolio/summarytrees@htmlwidget")
```

## Examples


```{r "example01", eval = F}
# devtools::install_github("timelyportfolio/summarytrees@htmlwidget")

library(summarytrees)
data(dmoz)

#use example from vignette
K <- 100
g <- greedy(node = dmoz[, "node"],
           parent = dmoz[, "parent"],
           weight = dmoz[, "weight"],
           label = dmoz[, "label"],
           K = K)

# Prepare the summary trees for visualization:
json <- prepare.vis(tree.list = g$summary.trees,
                    labels = g$data[, "label"],
                    tree = g$tree,
                    legend.width = 150,
                    node.width = 225,
                    node.height = 14,
                    units = "# of URLs",
                    print.weights = TRUE,
                    legend.color = "lightsteelblue",
                    color.level = 3)
summarytrees_htmlwidget(json)
```

```{r "example01_run", echo = F}
# devtools::install_github("timelyportfolio/summarytrees@htmlwidget")

library(summarytrees)
data(dmoz)

#use example from vignette
K <- 100
g <- greedy(node = dmoz[, "node"],
           parent = dmoz[, "parent"],
           weight = dmoz[, "weight"],
           label = dmoz[, "label"],
           K = K)

# Prepare the summary trees for visualization:
json <- prepare.vis(tree.list = g$summary.trees,
                    labels = g$data[, "label"],
                    tree = g$tree,
                    legend.width = 150,
                    node.width = 225,
                    node.height = 14,
                    units = "# of URLs",
                    print.weights = TRUE,
                    legend.color = "lightsteelblue",
                    color.level = 3)
summarytrees_htmlwidget(json, height = 500, width = 600 ) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 680
      , height = 580
      , overflow = "hidden"
    )
  )
```


## Thanks

Thanks [Kenneth Shirley](http://www.kennyshirley.com/) for the research and implementation of `summarytrees`.

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

