---
title: "Week25 - sweetAlertR"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## htmlwidgets News This Week

I can't keep up with all that is happening in widget-world, so for all the newest and updated `htmlwidgets`, just do this [Github search](https://github.com/search?o=desc&q=htmlwidgets&s=indexed&type=Code), and you'll quickly get up to date.

## This Week's Widget - `sweetalertR`

[SweetAlert](t4t5.github.io/sweetalert) by [Tristan Edwards](http://tristanedwards.me/) is a beautiful, dependency-free, and really popular replacement for JavaScript's ugly and crude  `alert`, so in my mind perfect for an `htmlwidget`.  [`sweetalertR`](https://github.com/timelyportfolio/sweetalertR) offers the beauty and power of SweetAlert to R in `htmlwidget` form.


## Quick Installation

`sweetalertR` is not yet on CRAN and won't be unless someone asks, so for now please install with `devtools::install_github`. 

```{r eval = F}
devtools::install_github("timelyportfolio/sweetalertR")
```

## SweetAlert Example Recreated in `rmarkdown`

In the [demo page](http://t4t5.github.io/sweetalert) for SweetAlert, the author Tristan Edwards demonstrates both his HTML/CSS/JS prowress and documents SweetAlert's functionality.  Let's recreate the example fully in `rmarkdown` with `sweetalertR`.  **Please note Tristan deserves all the credit for this code**.

<iframe src = "http://timelyportfolio.github.io/buildingwidgets/week25/sweetalert_examples.html" style = "height:600px;width:85%;"></iframe>



## Thanks

Thanks so much

- [Tristan Edwards](http://tristanedwards.me/) for [`SweetAlert`](https://github.com/t4t5/sweetalert)
- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

