---
title: "Week23 - formattable"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## htmlwidgets News This Week

Avid readers already know, but for all the newest and updated `htmlwidgets`, just do this [Github search](https://github.com/search?o=desc&q=htmlwidgets&s=indexed&type=Code), and you'll quickly get up to date.

## This Week's Widget - `formattable`

This week's widget [`formattable`](https://github.com/renkun-ken/formattable) is unique from all of the other widgets so far.  One of my favorite R developers [\@renkun-ken](https://github.com/renkun-ken) (also created `rlist` and `pipeR`) made this really nice package `formattable` to add styling and Excel-like conditional formatting to tables in `Rmarkdown`.  However, its design limited its use to `Rmd` contexts.  Of course, the magic of `htmlwidgets` can solve this, so we created a new function `as.htmlwidget` to convert `formattable` objects and added some logic to auto-convert.  Kun Ren deserves all the credit for the awesomeness of `formattable`.  I simply assisted in the `htmlwidget` conversion and testing.


## Quick Installation

`formattable` is not yet on CRAN, so for now please install with `devtools::install_github`. 

```{r eval = F}
# should move to master soon
devtools::install_github("renkun-ken/formattable")
```

## Vignette

I tried to discuss in the vignette added to the package.  I'll embed it as an iframe below.

<iframe src = "http://timelyportfolio.github.io/buildingwidgets/week23/formattable_vignette.html" style = "height:800px; width:90%;"></iframe>

## Thanks

Thanks so much

- [\@renkun-ken](http://renkun.me) for `formattable`, `rlist`, and `pipeR`
- authors of [`markdown`](http://cran.r-project.org/web/packages/markdown/index.html)
- Yihui Xie for [`knitr`](http://yihui.name/knitr/)
- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

