---
title: "Week48 - stmCorrViz"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)
```

## This Week's Widget - `stmCorrViz` | htmlwidgetized stmCorrViz 

Way back in [week 13](www.buildingwidgets.com/blog/2015/4/2/week-13-interactive-stm), I html-widgetized [`stmBrowser`](https://github.com/mroberts/stmBrowser) which pairs with [`stm`](http://structuraltopicmodel.com/) and [`stmCorrViz`](https://github.com/AntonioCoppola/stmCorrViz).  [Scott Jacobs](https://www.linkedin.com/in/scottpatrickjacobs) very kindly asked if I could also html-widgetize `stmCorrViz`, and I said yes.  While at it, I also fixed a couple bugs in `stmBrowser`.

**Please note that any screwups or malfunctions are mine and not the original authors.**

### Quick Installation

The `htmlwidget` version of `stmBrowser` currently exists as a fork with a different branch on Github.  For now please install with `devtools::install_github`.  Also, you'll need [`stm`](http://cran.r-project.org/web/packages/stm/index.html) from CRAN.

```{r eval = F}
#install.packages("stm")
#devtools::install_github("timelyportfolio/stmBrowser@htmlwidget")
devtools::install_github("timelyportfolio/stmCorrViz@htmlwidget")
```


### Example in a Presentation

Remember I know basically nothing about Structural Topic Models (stm), so I'll just use one of the examples from `?stm::stm` and then slap both the `stmCorrViz` and `stmBrowser` in a `rmarkdown` presentation to prove that it really is an `htmlwidget`.  This demonstrates one of the many benefits of `htmlwidgets`.

<iframe src = "http://timelyportfolio.github.io/buildingwidgets/week48/stm_presentation.html" style = "height: 600px; width:100%; border: " seamless></iframe>


## Thanks

Thanks so much for all the work by

- [Antonio Coppola](https://github.com/AntonioCoppola), [Michael Freeman](http://mfviz.com), [Jason Chuang](http://jason.chuang.info), [Molly Roberts](http://margaretroberts.net), [Brandon Stewart](http://scholar.harvard.edu/bstewart) and [Dustin Tingley](http://scholar.harvard.edu/dtingley) for letting me barge in on their great work and mess it up in who knows how many ways
- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R`

