---
title: "Week04 - Interactive and Configurable Parallel Coordinates"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
---

```{r "setup",echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
  #, results = 'asis'
)

library(pipeR)
```

## htmlwidgets News This Week
@kbroman looks like he is exploring converting his `qtlcharts` to `htmlwidgets`.

-[\@kbroman](https://github.com/kbroman/testjs) - test `htmlwidgets` conversion of the scatterplot from his [`qtlcharts`](http://kbroman.org/qtlcharts/)

@rich-iannone

[\@rich-iannone] - [`DiagrammeR`](https://github.com/rich-iannone/DiagrammeR) hits version 0.4 with full shiny support and `graphviz`, gets a [webpage](http://rich-iannone.github.io/DiagrammeR/) (thanks JJ Allaire), and becomes the first `htmlwidget` to surpass 100 stars on Github.

## This Week's Widget - `parcoords` for Interactive d3.js Parallel Coordinates

Unfortunately, this week's widget is still alpha.


## What Does it Do?
The [visualizations from `networkD3`](http://christophergandrud.github.io/networkD3/) offer some really beautiful chart options for all sorts of network data. 

Link to parallel coordinates [paper](http://www.cs.tufts.edu/~remco/publications/2014/InfoVis2014-JND.pdf)

### Two Lines of Code - Existing Niceness

```{r "example01",eval = F}
library(networkD3)

# use this simple hclust example from
#  http://rpubs.com/gaston/dendrograms
hc = hclust(dist(mtcars))

treeNetwork( 
  as.treeNetwork(hc, "mtcars")
)
```
```{r "example01_run",echo = F}
library(networkD3)

hc = hclust(dist(mtcars))

treeNetwork( 
  as.treeNetwork(hc, "mtcars")
  , width = 750, height =750
) %>>%
      htmlwidgets::as.iframe(
        file="example01.html"
        , libdir = "lib"
        , selfcontained=F
        , width = 850
        , height =900
        , overflow = "hidden"
      )
```



### More Layouts with hierNetwork

Since this is not yet officially part of `networkD3`, to reproduce you will need to use `devtools::install_github`.

```{r eval = F}
devtools::install_github("timelyportfolio/networkD3@feature/d3.chart.layout")
```

```{r "example02", eval = F}
library(htmltools) 

tagList(
  lapply(
    c("tree.cartesian"
      ,"tree.radial"
      ,"cluster.cartesian"
      ,"cluster.radial"
    )
    ,function(chartType){
      hierNetwork( as.treeNetwork(hc), type = chartType, zoomable = T, collapsible = T )
    }
  )
)
```

```{r "example02_run", echo = F}
library(htmltools) 

i = 1
tagList(
  lapply(
    c("tree.cartesian"
      ,"tree.radial"
      ,"cluster.cartesian"
      ,"cluster.radial"
    )
    ,function(chartType){
      hn <- hierNetwork( as.treeNetwork(hc), type = chartType, zoomable = T, collapsible = T, margin = list(left=100), height = 750, width = 750 ) %>>%
        htmlwidgets::as.iframe(
        file=paste0("example02_",i,".html")
        , libdir = "lib"
        , selfcontained=F
        , width = 850
        , height =900
        , overflow = "hidden"
      )
      i <<- i + 1
      return(list(
         tags$h4(chartType)
         ,hn
        )
      )
    }
  )
)
```

For those of you wondering if there is a difference between `cluster` and `tree`, collapse some of the nodes.  If you still don't see it, `cluster` extends to the outermost level ( much like `hang` in `R` when plotting dendrograms ) while `tree` does not.  Also, I hope you enjoyed the pan/zoom functionality, which is enabled by `zoomable=T` and `collapsible=T`.  I admit it is hard to know that this interaction is available.

### Glitzier Layouts

Now on to some "glitzier" layouts that unlike the above layouts require a size or value.  These layouts in `d3.chart.layout` hold the most promise, but also have the most room for improvement.  I'll increase the complexity of the code by visualizing a `rpart` on the `diamonds` dataset from `ggplot2`.  If you only care about the visualizations, you can safely skip the data preparation below.  If you **care about the `rpart`/`partykit` piece** of this, **please let me know** since I am very interested in testing, improving, and iterating this, but need some help.


```{r "example02_data", results = F}
library(rpart)
library(partykit)
library(rlist)
library(pipeR)

#set up a little rpart as an example
# using data from ggplot2 diamonds dataset

data("diamonds",package="ggplot2")

rp <- rpart(
  price ~ carat + cut + color + clarity + depth + table + x + y + z
  ,method = "anova"
  ,data = diamonds
  ,control = rpart.control(minsplit = 2)
)

rpk <- as.party(rp)

## get meta information
rpk.text <- capture.output( print(rpk) ) %>>%
  ( .[grep( x = ., pattern = "(\\[)([0-9]*)(\\])")] ) %>>%
  strsplit( "[\\[\\|\\]]" , perl = T) %>>%
  list.map(
    tail(.,2) %>>%
      (
        data.frame(
          "id" = as.numeric(.[1])
          , description = .[2]
          , stringsAsFactors = F )
      )
  ) %>>% list.stack

# binding the node names from rpk with more of the relevant meta data from rp
# i don't think that partykit imports this automatically for the inner nodes, so i did it manually
rpk.text <- cbind(rpk.text, rp$frame)

# rounding the mean DV value
rpk.text$yval <- round(rpk.text$yval, 2)

# terminal nodes have descriptive stats in their names, so I stripped these out
# so the final plot wouldn't have duplicate data
rpk.text$description <- sapply(strsplit(rpk.text[,2], ":"), "[", 1)


dat = rapply(rpk$node,unclass,how="replace")
#fill in information at the root level for now
#that might be nice to provide to our interactive graph
dat$info = rapply(
  unclass(rpk$data)[-1]
  ,function(l){
    l = unclass(l)
    if( class(l) %in% c("terms","formula","call")) {
      l = paste0(as.character(l)[-1],collapse=as.character(l)[1])
    }          
    attributes(l) <- NULL
    return(l)
  }
  ,how="replace"
)

dat = jsonlite::toJSON(
  dat
  ,auto_unbox = T
)

# replace kids with children to ease d3
dat = gsub( x=dat, pattern = "kids", replacement="children")
# change id to node to ease d3; will replace with name later
dat = gsub ( x=dat, pattern = '"id":([0-9]*)', replacement = '"name":"node\\1","size":nodesize\\1' )
# calling the root node by the dataset name, but it might make more sense to call it
# "root" so that the code can be generalized
dat = sub (x = dat, pattern = "node1", replacement = "diamonds")
# replacing the node names from node1, node2, etc., with the extracted node names and metadata from
# rpk.text, and rp$table. 
for (i in 2:nrow(rpk.text)) {
  dat = sub (
    x = dat
    , pattern = paste("node", i, sep = "")
    , replacement = paste(
      rpk.text[i,2]
      , ", mean = ", rpk.text[i,7]
      , ", n = ", rpk.text[i,4]
      , sep = ""
    )
    , fixed = T
  )
  dat = sub (
    x = dat
    , pattern = paste("nodesize", i, sep = "")
    , replacement = rpk.text[i,4]
    , fixed = T
  )
}

# replace size of root or node1
dat = sub (
  x = dat
  , pattern = "nodesize1"
  , replacement = rpk.text[1,4]
  , fixed = T
)
```


You can probably see **why data conversion might possibly be the most valuable part of an `htmlwidget`**.  Now let's visualize with our other layouts from `d3.chart.layout`.  I intentionally did not use the common `flare.json` dataset that drives nearly all examples of this type.  If you want to see these layouts with this dataset, check out the `d3.chart.layout` [examples](http://bl.ocks.org/bansaghi).  Also, I intentionally left these charts in a raw form, so please don't dismiss these because the labelling isn't right.


```{r "example_03", eval= F}
hN <- hierNetwork( jsonlite::fromJSON(dat), zoomable = T, collapsible = T )
# fromJSON does not translate well so manual override
hN$x$root = dat
lapply(
  c("pack.nested"
    ,"pack.flattened"
    ,"partition.arc"
    ,"partition.rectangle"
    ,"treemap"
  )
  ,function(chartType){
    hN$x$options$type = chartType
    return(hN) 
  }
)
```

```{r "example03_run", echo = F}
hN <- hierNetwork( jsonlite::fromJSON(dat), zoomable = T, collapsible = T )
# fromJSON does not translate well so manual override
hN$x$root = dat
i = 1
tagList(
  lapply(
    c("pack.nested"
        ,"pack.flattened"
        ,"partition.arc"
        ,"partition.rectangle"
        ,"treemap"
    )
    ,function(chartType){
      hN$x$options$type = chartType

      hN %>>%
        htmlwidgets::as.iframe(
        file=paste0("example03_",i,".html")
        , libdir = "lib"
        , selfcontained=F
        , width = 850
        , height =900
        , overflow = "hidden"
      ) -> iframe
      i <<- i + 1
      return(list(
         tags$h4(chartType)
         ,iframe
        )
      )
    }
  )
)

```

### Direction I'd Like to Head

As a demonstration of what we can achieve with these type layouts with a little more work, please see these examples.


1. `pack.nested` from Bill White [example](http://www.billdwhite.com/wordpress/2013/10/28/circle-pack-label-truncation/)

<iframe src = "http://bl.ocks.org/billdwhite/raw/7207695/" overflow = "hidden" height = 700 width = 900 "seamless"></iframe>

2.  `pack.flattened` from Mike Bostock [example](http://bl.ocks.org/mbostock/7882658)

<iframe src = "http://bl.ocks.org/mbostock/raw/7882658" overflow = "hidden" height = 700 width = 900 "seamless"></iframe>

3.  `partition.arc` from Kerry Rodden [example](http://bl.ocks.org/kerryrodden/7090426)

<iframe src = "http://bl.ocks.org/kerryrodden/raw/7090426/" overflow = "hidden" height = 700 width = 900 "seamless"></iframe>

4.  `treemap` from Solomon Kahn on the U.S. budget [example](http://solomonkahn.com/us_budget/)



## Thanks

Thanks so much for all the work by

- [Mike Bostock](http://bost.ocks.org/mike/)
- [Kai Chang](https://github.com/syntagmatic)

