---
title: "Week46 - vegaliteR"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set( 
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## This Week's Widget - vegaliteR

When [`vega`](https://github.com/vega/vega) was first released, I spent a lot of time exploring its potential.  Then [`ggvis`](https://github.com/rstudio/ggvis) sort of eliminated the need to learn the internals of `vega`.  Then `ggvis` stalled, and `vega` got lots of new features (mainly [frp interactivity](http://idl.cs.washington.edu/files/2015-ReactiveVega-InfoVis.pdf)), and the team at University of Washington Interactive Data Lab went hyper-productive releasing the vega derivatives [`voyager`](https://github.com/vega/voyager), [`polestar`](https://github.com/vega/polestar), [`compass`](https://github.com/vega/compass), and [`vega-lite`](https://github.com/vega/vega-lite).  After listening to [Arvind Satyanarayan](http://arvindsatya.com/) on the [PolicyVis podcast](http://policyviz.com/episode-23-arvind-satyanarayn/), I decided that a `vega-lite` htmlwidget would help get me caught up.

This week's widget [`vegaliteR`](https://github.com/timelyportfolio/vegaliteR) is the result of that exercise.  Currently, it doesn't come close to fulfilling the checklist for a good `htmlwidget`, but this very literal interpretation will require learning the easy `vega-lite` schema.  With some iteration, I hope to make it much more `R`-like.

## Installation

This is not on CRAN, so to install we will need some help from `devtools::install_github`.

```{r eval = FALSE}
devtools::install_github("timelyportfolio/vegaliteR")
```

## data.frame into a list of lists

Most good `htmlwidgets` handle this bit for us, but like I said in the introduction, this first release is a very literal interpretation, so the code might look a little strange with nested lists and data as an array of objects (or `dataframe="rows"` in `jsonlite`).  For those familiar with `rCharts` we deliberately avoided all this, since it can be a little bewildering for an `R` user.  However, there is certainly no harm in straddling the line between `R` and `JavaScript`.  Plus, learning some new techniques with [`purrr`](https://github.com/hadley/purrr) and [`rlist`](https://github.com/renkun-ken/rlist) might even be fun.

Let's start by looking at 3 ways in `R` we can make a `data.frame` into a `list` of `lists` or an `array` of `objects`.

### with apply

`apply` seems easy, but doesn't really work since `apply` coerces to `array` or `matrix`, so we lose `class`.  We can make it work, but I'm lazy, so let's skip it.

```{r eval = FALSE}
unname(apply(mtcars, MARGIN=1, as.list))
```

### with rlist

[`rlist`](http://renkun.me/blog/2014/08/06/introducing-rlist-0.3.html) has lots of `list` helpers, including `list.parse`.

```{r eval = FALSE}
library(rlist)

# without rownames
unname(list.parse(mtcars))

# with rownames
unname(list.parse(
  data.frame(name=rownames(mtcars),mtcars,stringsAsFactors = FALSE)
))
```

### with purrr

[`purrr`](https://github.com/hadley/purrr) provides `by_row`.  Let's see how we can use it.

```{r eval = FALSE}
library(purrr)

# without rownames
by_row(mtcars,as.list)$.out

# with rownames
by_row(
  data.frame(name=rownames(mtcars),mtcars,stringsAsFactors = FALSE),
  as.list
)$.out
)
```

## Examples

Now that we know three different ways of converting a `data.frame` to a `list` of `lists`, let's make some `vega` charts.  To make it easy, we'll store our converted data.

```{r}
# remove . from colnames
colnames(swiss) <- gsub(
  x = colnames(swiss),
  pattern = "\\.",
  replacement = ""
)
swiss_list <- unname(rlist::list.parse(
  data.frame(village=rownames(swiss),swiss,stringsAsFactors = FALSE)
))
```

### Scatter Plot

```{r "example01", eval = F}
#devtools::install_github("timelyportfolio/vegaliteR")
library(vegaliteR)

vegalite(
  list(
    data = list(values = swiss_list),
    marktype = "point",
    encoding = list(
      x = list(field = "Fertility", type = "Q"),
      y = list(field = "InfantMortality", type = "Q")
    )
  )
)
```

```{r "example01_run", echo = F}
#devtools::install_github("timelyportfolio/vegaliteR")
library(vegaliteR)

vegalite(
  list(
    data = list(values = swiss_list),
    marktype = "point",
    encoding = list(
      x = list(field = "Fertility", type = "Q"),
      y = list(field = "InfantMortality", type = "Q")
    )
  )
,height = 300, width = 500) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 650
      , height = 450
      , overflow = "hidden"
    )
  )
```

`vega-lite` also makes aggregation easy.

```{r "example02", eval = F}
#devtools::install_github("timelyportfolio/vegaliteR")
library(vegaliteR)
library(pipeR)

lapply(
  swiss_list,
  function(x){
    c(
      majority_catholic = if(x[["Catholic"]]>50) "yes" else "no",
      x
    )
  }
) %>>%
  (
    list(
      data = list(values = .),
      marktype = "square",
      encoding = list(
        y = list(field="Fertility",type="Q",aggregate="mean"),
        x = list(field="majority_catholic",type="nominal"),
        color = list(field="majority_catholic",type="nominal")
      ),
      config = list(
        largeBandWidth = 60
      )
    )
  ) %>>%
  vegalite()
```

```{r "example02_run", echo = F}
#devtools::install_github("timelyportfolio/vegaliteR")
library(vegaliteR)
library(pipeR)

lapply(
  swiss_list,
  function(x){
    c(
      majority_catholic = if(x[["Catholic"]]>50) "yes" else "no",
      x
    )
  }
) %>>%
  (
    list(
      data = list(values = .),
      marktype = "square",
      encoding = list(
        y = list(field="Fertility",type="Q",aggregate="mean"),
        x = list(field="majority_catholic",type="nominal"),
        color = list(field="majority_catholic",type="nominal")
      ),
      config = list(
        largeBandWidth = 60
      )
    )
  ) %>>%
  vegalite(height = 300,width = 400) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example02.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 550
      , height = 450
      , overflow = "hidden"
    )
  )
```

### Line Chart

For a line chart, we simply change `marktype = "line"`.  As a not best-practice example, we can plot `Agriculture` as a function of `Education`.

```{r "example03", eval = F}
#devtools::install_github("timelyportfolio/vegaliteR")
library(vegaliteR)
library(pipeR)

swiss_list %>>%
  (
    list(
      data = list(values = .),
      marktype = "line",
      encoding = list(
        y = list(field="Agriculture",type="Q"),
        x = list(field="Education",type="Q")
      )
    )
  ) 
```

```{r "example03_run", echo = F}
#devtools::install_github("timelyportfolio/vegaliteR")
library(vegaliteR)
library(pipeR)

swiss_list %>>%
  (
    list(
      data = list(values = .),
      marktype = "line",
      encoding = list(
        y = list(field="Agriculture",type="Q"),
        x = list(field="Education",type="Q")
      )
    )
  ) %>>%
  vegalite(height = 300, width = 400) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example03.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 550
      , height = 450
      , overflow = "hidden"
    )
  )
```


### Bar Chart

We can easily make our aggregate scatter example into a bar chart by changing the `marktype`.

```{r "example04", eval = F}
#devtools::install_github("timelyportfolio/vegaliteR")
library(vegaliteR)
library(pipeR)

lapply(
  swiss_list,
  function(x){
    c(
      majority_catholic = if(x[["Catholic"]]>50) "yes" else "no",
      x
    )
  }
) %>>%
  (
    list(
      data = list(values = .),
      marktype = "bar",
      encoding = list(
        y = list(field="Fertility",type="Q",aggregate="mean"),
        x = list(field="majority_catholic",type="nominal"),
        color = list(field="majority_catholic",type="nominal")
      ),
      config = list(
        largeBandWidth = 60
      )
    )
  ) %>>%
  vegalite()
```

```{r "example04_run", echo = F}
library(vegaliteR)
library(pipeR)

lapply(
  swiss_list,
  function(x){
    c(
      majority_catholic = if(x[["Catholic"]]>50) "yes" else "no",
      x
    )
  }
) %>>%
  (
    list(
      data = list(values = .),
      marktype = "bar",
      encoding = list(
        y = list(field="Fertility",type="Q",aggregate="mean"),
        x = list(field="majority_catholic",type="nominal"),
        color = list(field="majority_catholic",type="nominal")
      ),
      config = list(
        largeBandWidth = 60
      )
    )
  ) %>>%
  vegalite(height = 300, width = 400) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example04.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 550
      , height = 450
      , overflow = "hidden"
    )
  )
```

### Stacked Bar Chart

For a facetted stacked bar, we can recreate the `vega-lite` [stacked bar](http://vega.github.io/vega-editor/?mode=vega-lite&spec=trellis_stacked_bar&renderer=svg).

```{r "example05", eval = F}
#devtools::install_github("timelyportfolio/vegaliteR")
library(vegaliteR)
library(pipeR)
library(purrr)

data(barley, package="lattice")

by_row(barley,as.list)$.out %>>%
  (
    list(
      data = list(values = .),
      marktype = "bar",
      encoding = list(
        x = list(field="yield",type="Q",aggregate="sum"),
        y = list(field="variety",type="nominal"),
        color = list(field="site",type="nominal")
      )
    )
  ) %>>%
  vegalite()
```

```{r "example05_run", echo = F}
library(vegaliteR)
library(pipeR)
library(purrr)

data(barley, package="lattice")

by_row(barley,as.list)$.out %>>%
  (
    list(
      data = list(values = .),
      marktype = "bar",
      encoding = list(
        x = list(field="yield",type="Q",aggregate="sum"),
        y = list(field="variety",type="nominal"),
        color = list(field="site",type="nominal")
      )
    )
  ) %>>%
  vegalite(height = 300, width = 400) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example05.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 550
      , height = 450
      , overflow = "hidden"
    )
  )
```

## More

That's probably enough for this post.  I'll try to make `vegaliteR` more `R`-like as I get time and the API stabilizes.  For more `vega-lite` examples, see the [vega-editor](http://vega.github.io/vega-editor/?mode=vega-lite).

Happy Thanksgiving to those in the US.

## Thanks

Thanks [University of Washington Interactive Data Lab](http://idl.cs.washington.edu/) for their hyper-productive efforts with `vega` and all its derivatives.

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

