---
title: "Week49 - d3radarR"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set( 
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## Bleeding Edge - crosstalk

If you want to get a preview of the exciting changes for `htmlwidgets`, be sure to check out `htmlwidgets` [issue 86](https://github.com/ramnathv/htmlwidgets/issues/86#issuecomment-159863989).  Thanks to Joe Cheng from RStudio for taking the lead on this.

## This Week's Widget - d3radarR

[Charles Allen](https://github.com/TennisVisuals) at [TennisVisuals](http://tennisvisuals.com/) blogging at [tennisvis.blogspot.com](http://tennisvis.blogspot.com) has been doing some really incredible visualization work for tennis.  When I saw his [reusable iteration](https://github.com/TennisVisuals/updating-radar-chart) of [Nadieh Bremer's](https://twitter.com/NadiehBremer) redesigned [radar chart](http://www.visualcinnamon.com/2015/10/different-look-d3-radar-chart.html) I instantly wanted it in `R`.  This week I offer an early and rough `htmlwidget` implementation in [`d3radarR`](https://github.com/timelyportfolio/d3radarR).  **I only wish I could more fully support all of the options and accessors included with the radar chart.**  Next year, once I wrap up a widget a week, I hope to revisit this.  As always, any feedback and polite requests motivate and inspire me.

## Installation

This is not on CRAN, so to install we will need some help from `devtools::install_github`.

```{r eval = FALSE}
devtools::install_github("timelyportfolio/d3radarR")
```

## Examples

Recreating the included examples is usually a good place to start.  Let's recreate the phone data radar chart.

```{r "example01", eval = F}
# devtools::install_github("timelyportfolio/d3radarR")

library(d3radarR)

# use data from demo
#   https://github.com/TennisVisuals/updating-radar-chart/blob/master/radarDemo.js
json_data = jsonlite::fromJSON(
'
  [  
    {  
      "key":"Nokia Smartphone",
      "values":[  
        {  "axis":"Battery Life", "value":0.26 }, {  "axis":"Brand", "value":0.10 },
        {  "axis":"Contract Cost", "value":0.30 }, {  "axis":"Design And Quality", "value":0.14 },
        {  "axis":"Have Internet Connectivity", "value":0.22 }, {  "axis":"Large Screen", "value":0.04 },
        {  "axis":"Price Of Device", "value":0.41 }, {  "axis":"To Be A Smartphone", "value":0.30 }
        ]
    },
    {  
      "key":"Samsung",
      "values":[  
        {  "axis":"Battery Life", "value":0.27 }, {  "axis":"Brand", "value":0.16 },
        {  "axis":"Contract Cost", "value":0.35 }, {  "axis":"Design And Quality", "value":0.13 },
        {  "axis":"Have Internet Connectivity", "value":0.20 }, {  "axis":"Large Screen", "value":0.13 },
        {  "axis":"Price Of Device", "value":0.35 }, {  "axis":"To Be A Smartphone", "value":0.38 }
        ]
    },
    {  
      "key":"iPhone",
      "values":[  
        {  "axis":"Battery Life", "value":0.22 }, {  "axis":"Brand", "value":0.28 },
        {  "axis":"Contract Cost", "value":0.29 }, {  "axis":"Design And Quality", "value":0.17 },
        {  "axis":"Have Internet Connectivity", "value":0.22 }, {  "axis":"Large Screen", "value":0.02 },
        {  "axis":"Price Of Device", "value":0.21 }, {  "axis":"To Be A Smartphone", "value":0.50 }
        ]
    }
  ]
',
  simplifyDataFrame = FALSE
)

d3radar(json_data)
```

```{r "example01_run", echo = F}
library(d3radarR)

# use data from demo
#   https://github.com/TennisVisuals/updating-radar-chart/blob/master/radarDemo.js
json_data = jsonlite::fromJSON(
'
  [  
    {  
      "key":"Nokia Smartphone",
      "values":[  
        {  "axis":"Battery Life", "value":0.26 }, {  "axis":"Brand", "value":0.10 },
        {  "axis":"Contract Cost", "value":0.30 }, {  "axis":"Design And Quality", "value":0.14 },
        {  "axis":"Have Internet Connectivity", "value":0.22 }, {  "axis":"Large Screen", "value":0.04 },
        {  "axis":"Price Of Device", "value":0.41 }, {  "axis":"To Be A Smartphone", "value":0.30 }
        ]
    },
    {  
      "key":"Samsung",
      "values":[  
        {  "axis":"Battery Life", "value":0.27 }, {  "axis":"Brand", "value":0.16 },
        {  "axis":"Contract Cost", "value":0.35 }, {  "axis":"Design And Quality", "value":0.13 },
        {  "axis":"Have Internet Connectivity", "value":0.20 }, {  "axis":"Large Screen", "value":0.13 },
        {  "axis":"Price Of Device", "value":0.35 }, {  "axis":"To Be A Smartphone", "value":0.38 }
        ]
    },
    {  
      "key":"iPhone",
      "values":[  
        {  "axis":"Battery Life", "value":0.22 }, {  "axis":"Brand", "value":0.28 },
        {  "axis":"Contract Cost", "value":0.29 }, {  "axis":"Design And Quality", "value":0.17 },
        {  "axis":"Have Internet Connectivity", "value":0.22 }, {  "axis":"Large Screen", "value":0.02 },
        {  "axis":"Price Of Device", "value":0.21 }, {  "axis":"To Be A Smartphone", "value":0.50 }
        ]
    }
  ]
',
  simplifyDataFrame = FALSE
)

d3radar(json_data, height = 400, width = 400) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 550
      , height = 550
      , overflow = "hidden"
    )
  )
```

You might notice the data format is not very `R`-like.  One of the most immediate needs for this implementation would be a data transformer for `R`.  I unfortunately ran out of time this week, so we might need to learn a little `dplyr` and `purrr` to work with `data.frames`.  Let's look at a `mtcars` as a radar.


```{r "example02", eval = F}
# devtools::install_github("timelyportfolio/d3radarR")

# use example from ?stars
library(d3radarR)
library(scales)
library(dplyr)
library(purrr)

# this is ugly
#  should probably include some conversion helpers
mtcars %>%
  add_rownames( var = "key" ) %>%
  mutate_each(funs(rescale), -key) %>%
  #limit to 8 rows for manageability
  head(8) %>%
  by_row(
    function(row){
      list(
        key = row[["key"]],
        values = lapply(
          names(row)[-match("key",names(row))],
          function(nm){
            list(
              "axis" = nm,
              "value" = row[[nm]]
            )
          }
        )
      )
    }
  ) %>%
  {
    d3radar( .$.out )
  }

```

```{r "example02_run", echo = F}
# use example from ?stars
library(d3radarR)
library(scales)
library(dplyr)
library(purrr)

# this is ugly
#  should probably include some conversion helpers
mtcars %>%
  add_rownames( var = "key" ) %>%
  mutate_each(funs(rescale), -key) %>%
  #limit to 8 rows for manageability
  head(8) %>%
  by_row(
    function(row){
      list(
        key = row[["key"]],
        values = lapply(
          names(row)[-match("key",names(row))],
          function(nm){
            list(
              "axis" = nm,
              "value" = row[[nm]]
            )
          }
        )
      )
    }
  ) %>%
  {
    d3radar( .$.out, height = 400, width = 600 )
  } %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example02.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 750
      , height = 550
      , overflow = "hidden"
    )
  )
```

I hear that a certain team that I might like might be playing an important game soon.  Let's looks at a radar chart using [ESPN efficiency ratings](http://espn.go.com/college-football/statistics/teamratings/_/tab/efficiency).  Here is where colors would be really helpful, but I just didn't have the time.

```{r "example03", eval = F}
# devtools::install_github("timelyportfolio/d3radarR")

library(d3radarR)

# source : http://espn.go.com/college-football/statistics/teamratings/_/tab/efficiency
espn_efficiency <- list(
  list(
    key = "Alabama",
    values = list(
      list(axis = "offense", value = 70.9/100),
      list(axis = "defense", value = 92.9/100),
      list(axis = "specialtms", value = 58.3/100)
    )
  ),
  list(
    key = "Oklahoma",
    values = list(
      list(axis = "offense", value = 83.3/100),
      list(axis = "defense", value = 83.1/100),
      list(axis = "specialtms", value =42.6/100)
    )
  ),
  list(
    key = "Clemson",
    values = list(
      list(axis = "offense", value = 81.3/100),
      list(axis = "defense", value = 87.1/100),
      list(axis = "specialtms", value = 32.4/100)
    )
  ),
  list(
    key = "Michigan State",
    values = list(
      list(axis = "offense", value = 70.5/100),
      list(axis = "defense", value = 73.2/100),
      list(axis = "specialtms", value = 39.0/100)
    )
  )  
)

d3radar( espn_efficiency )
```

```{r "example03_run", echo = F}
library(d3radarR)

# source : http://espn.go.com/college-football/statistics/teamratings/_/tab/efficiency
espn_efficiency <- list(
  list(
    key = "Alabama",
    values = list(
      list(axis = "offense", value = 70.9/100),
      list(axis = "defense", value = 92.9/100),
      list(axis = "specialtms", value = 58.3/100)
    )
  ),
  list(
    key = "Oklahoma",
    values = list(
      list(axis = "offense", value = 83.3/100),
      list(axis = "defense", value = 83.1/100),
      list(axis = "specialtms", value =42.6/100)
    )
  ),
  list(
    key = "Clemson",
    values = list(
      list(axis = "offense", value = 81.3/100),
      list(axis = "defense", value = 87.1/100),
      list(axis = "specialtms", value = 32.4/100)
    )
  ),
  list(
    key = "Michigan State",
    values = list(
      list(axis = "offense", value = 70.5/100),
      list(axis = "defense", value = 73.2/100),
      list(axis = "specialtms", value = 39.0/100)
    )
  )  
)

d3radar( espn_efficiency, height = 400, width = 600 ) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example03.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 750
      , height = 550
      , overflow = "hidden"
    )
  )
```


## Thanks

Thanks [Charles Allen](https://github.com/TennisVisuals) at [TennisVisuals](http://tennisvisuals.com/) for his great work and in particular his responsiveness to my [pestering](https://github.com/TennisVisuals/updating-radar-chart/issues?q=is%3Aissue+is%3Aclosed) on his updating radar chart.

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

