---
title: "Week27 - d3hiveR"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## htmlwidgets News This Week

Yay, it appears a **gallery is now in the works**.  Until its release though, for all the newest and updated `htmlwidgets`, just do this [Github search](https://github.com/search?o=desc&q=htmlwidgets&s=indexed&type=Code), and you'll quickly get up to date.

## This Week's Widget - `d3hiveR`

[Martin Krzywinski](http://mkweb.bcgsc.ca/) makes beautiful things.  Just have a look at his site for [hive plots](http://www.hiveplot.net) which he created back in 2011.

<blockquote>
<p class="citation"><em>Published in BRIEFINGS IN BIOINFORMATICS</em><br> Krzywinski M, Birol I, Jones S, Marra M (2011). <a href="http://bib.oxfordjournals.org/content/early/2011/12/09/bib.bbr069.short?rss=1">Hive Plots — Rational Approach to Visualizing Networks</a>. Briefings in Bioinformatics (early access 9 December 2011, doi: 10.1093/bib/bbr069). (<a href="hiveplot.enw">download citation</a>)</p>
</blockquote>
Hive plots are beautiful on their own, but they also serve a purpose to help visualize complex networks, or tame the network hairball (for more, see [Lynn Cherny's presentation](http://www.slideshare.net/OReillyStrata/visualizing-networks-beyond-the-hairball)).  `R` already has the very good, well-documented [`HiveR`](http://cran.r-project.org/web/packages/HiveR/index.html) from [Bryan Hanson](https://github.com/bryanhanson).  So, this week's [`d3hiveR`](https://github.com/timelyportfolio/d3hiveR) has a strong headstart on the `R` side.  On the JavaScript side, we have a couple examples.

- [Hive Plots](http://bost.ocks.org/mike/hive/) by Mike Bostock
- [Hive Plot Links](http://bl.ocks.org/mbostock/2066415) by Mike Bostock
- [Hive Plot Demo Rework](http://bl.ocks.org/RichMorin/2117857) by [Rich Morin](http://www.cfcl.com/rdm/)
- [Hive Plot for Student Systems](http://bl.ocks.org/saraquigley/2346962) by Sara Quiqley
- [Hive Plot Areas](http://bl.ocks.org/mbostock/2066421) by Mike Bostock

Thanks to all of the above for all their work on hive plots.

## More Unfinished Than Any Other

All of the widgets created here should be considered unfinished, but this one is particulary undone, and the gaps are obvious.  I would love some help.

I just couldn't but be distracted by the thought of interactive [`tmap`](https://github.com/mtennekes/tmap) by Martijn Tennekes.  See this [issue](https://github.com/mtennekes/tmap/issues/20) to follow along or offer input.

## Quick Installation

`d3hiveR` is not yet on CRAN and won't be unless someone asks, so for now please install with `devtools::install_github`. 

```{r eval = F}
devtools::install_github("timelyportfolio/d3hiveR")
```


## Examples

`d3hiveR` leverages `HiveR`, so you can just feed it a `HivePlotData` object.

```{r "example01", eval = FALSE}
# devtools::install_github("timelyportfolio/d3hiveR")
library(HiveR)
library(d3hiveR)

d3hive( ranHiveData( nx = 5, ne = 1000, ns = c(1,3,8) ) )
```

```{r "example01_run", eval = TRUE, echo = FALSE}
# devtools::install_github("timelyportfolio/d3hiveR")
library(HiveR)
library(d3hiveR)

d3hive( ranHiveData( nx = 5, ne = 1000, ns = c(1,3,8) ) ) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 800
      , height = 600
      , overflow = "hidden"
   ))
```

## Thanks

Thanks so much to Martin Krzywinski, Bryan Hanson, Mike Bostock, Rich Morin, and Sara Quiqley all cited above.

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

