---
title: "Week42 - adjacency matrix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set( 
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## This Week's Widget - adjacency matrix

This beautiful little [adjacency matrix](http://bost.ocks.org/mike/miserables/) by Mike Bostock inspired this Github [issue/request](https://github.com/timelyportfolio/buildingwidgets/issues/64).  At the beginning of the week, I started banging away on a `htmlwidget` extension to [`networkD3`](https://github.com/christophergandrud/networkD3) using Elijah Meeks' [d3.layout.adjacencyMatrix](https://github.com/emeeks/d3.layout.adjacencyMatrix).  The deeper I got the more I realized that an adjacency matrix really is just a subset of a heatmap, so then I started replicating [Matthew Lincoln](http://matthewlincoln.net/)'s [fine `Shiny`/`ggplot2` adjacency matrix](http://matthewlincoln.net/2014/12/20/adjacency-matrix-plots-with-r-and-ggplot2.html) with RStudio/Joe Cheng's [`d3heatmap`](https://github.com/rstudio/d3heatmap).  This [bl.ock visualizing 2015 NCAA football](http://bl.ocks.org/timelyportfolio/0ae64825cf9b3278faf9) is one result.

As the end of this week nears, I don't have a whole lot in terms of a finished `htmlwidget`, but I do have some prototype code and some examples.

## Installation

This is not on CRAN and not even in the `master` branch, so to install we will need some help from `devtools::install_github`.

```{r eval = FALSE}
devtools::install_github("timelyportfolio/networkD3@feature/adjacency")
```

For `d3heatmap`, you can just `install.packages("d3heatmap")`.


## Examples

Let's start small with the [bridges of Koenigsberg](https://en.wikipedia.org/wiki/Seven_Bridges_of_K%C3%B6nigsberg).

```{r "example01", eval = F}
#devtools::install_github("timelyportfolio/networkD3@feature/adjacency")

library(networkD3)
library(igraph)

data(Koenigsberg, package="igraphdata")
koen_df <- get.data.frame(upgrade_graph(Koenigsberg), what="both")
koen_df$edges$from <- match(koen_df$edges$from,koen_df$vertices$name) - 1
koen_df$edges$to <- match(koen_df$edges$to,koen_df$vertices$name) - 1

adjacencyNetwork(
  Links = koen_df$edges,
  Nodes = koen_df$vertices,
  Source = "from",
  Target = "to",
  NodeID = "name",
  Group = "name",
  margin = list(left=150, top=150),
  colourScale = htmlwidgets::JS("d3.scale.category10()"),
  width = 500,
  height = 500
)
```

```{r "example01_run", echo = F}
#devtools::install_github("timelyportfolio/networkD3@feature/adjacency")

library(networkD3)
library(igraph)

data(Koenigsberg, package="igraphdata")
koen_df <- get.data.frame(upgrade_graph(Koenigsberg), what="both")
koen_df$edges$from <- match(koen_df$edges$from,koen_df$vertices$name) - 1
koen_df$edges$to <- match(koen_df$edges$to,koen_df$vertices$name) - 1

adjacencyNetwork(
  Links = koen_df$edges,
  Nodes = koen_df$vertices,
  Source = "from",
  Target = "to",
  NodeID = "name",
  Group = "name",
  margin = list(left=150, top=150),
  colourScale = htmlwidgets::JS("d3.scale.category10()"),
  width = 500,
  height = 500
) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 650
      , height = 650
      , overflow = "hidden"
   ))
```

Now we can compare with `d3heatmap`.

```{r "example02", eval = F}
library(d3heatmap)

d3heatmap(
  get.adjacency(
    upgrade_graph(Koenigsberg),
    sparse = FALSE
  ),
  colors = "Blues",
  dendrogram = "none",
  cexRow = 0.7, cexCol = 0.7
)
```

```{r "example02_run", echo = F}
library(d3heatmap)

d3heatmap(
  get.adjacency(
    upgrade_graph(Koenigsberg),
    sparse = FALSE
  ),
  colors = "Blues",
  dendrogram = "none",
  cexRow = 0.7, cexCol = 0.7,
  width = 500,
  height = 500
) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example02.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 650
      , height = 650
      , overflow = "hidden"
   ))
```

One more example with the `karate` data from `igraphdata`.

```{r "example03", eval = F}
library(networkD3)
library(igraph)

data(karate, package="igraphdata")
karate_df <- get.data.frame(upgrade_graph(karate), what="both")
karate_df$edges$from <- match(karate_df$edges$from,karate_df$vertices$name) - 1
karate_df$edges$to <- match(karate_df$edges$to,karate_df$vertices$name) - 1

adjacencyNetwork(
  Links = karate_df$edges,
  Nodes = karate_df$vertices,
  Source = "from",
  Target = "to",
  NodeID = "name",
  Group = "Faction",
  margin = list(left=150, top=150),
  width = 500,
  height = 500
)
```

```{r "example03_run", echo = F}
library(networkD3)
library(igraph)

data(karate, package="igraphdata")
karate_df <- get.data.frame(upgrade_graph(karate), what="both")
karate_df$edges$from <- match(karate_df$edges$from,karate_df$vertices$name) - 1
karate_df$edges$to <- match(karate_df$edges$to,karate_df$vertices$name) - 1

adjacencyNetwork(
  Links = karate_df$edges,
  Nodes = karate_df$vertices,
  Source = "from",
  Target = "to",
  NodeID = "name",
  Group = "Faction",
  margin = list(left=150, top=150),
  width = 500,
  height = 500
) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example03.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 650
      , height = 650
      , overflow = "hidden"
   ))
```

Again, let's do the same network but in `d3heatmap`, but we will include the dendrogram this time.

```{r "example04", eval = F}
library(d3heatmap)

d3heatmap(
  get.adjacency(
    upgrade_graph(karate),
    sparse = FALSE
  ),
  colors = "Blues",
  cexRow = 0.7, cexCol = 0.7
)
```

```{r "example04_run", echo = F}
library(d3heatmap)

d3heatmap(
  get.adjacency(
    upgrade_graph(karate),
    sparse = FALSE
  ),
  colors = "Blues",
  cexRow = 0.7, cexCol = 0.7,
  width = 500,
  height = 500
) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example04.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 650
      , height = 650
      , overflow = "hidden"
   ))
```

With a little sorting, our `karate` adjacency matrix can be more helpful.


```{r "example05", eval = F}
# community then by degree sorting
library(dplyr)
fc <- fastgreedy.community(karate)
karate_df <- get.data.frame(karate, what = "both")
karate_df$vertices <- data.frame(membership = unclass(membership(fc)), degree = degree(karate)) %>%
  mutate( name = rownames(.) ) %>%
  arrange( membership, desc( degree ) ) %>%
  select( name ) %>%
  as.vector %>%
  inner_join( karate_df$vertices )
karate_df$edges$from <- match(karate_df$edges$from,karate_df$vertices$name) - 1
karate_df$edges$to <- match(karate_df$edges$to,karate_df$vertices$name) - 1

adjacencyNetwork(
  Links = karate_df$edges,
  Nodes = karate_df$vertices[,c(2,1)],
  Source = "from",
  Target = "to",
  NodeID = "name",
  Group = "Faction",
  margin = list(left=100, top=100)
)
```

```{r "example05_run", echo = F}
# community then by degree sorting
library(dplyr)
fc <- fastgreedy.community(karate)
karate_df <- get.data.frame(karate, what = "both")
karate_df$vertices <- data.frame(membership = unclass(membership(fc)), degree = degree(karate)) %>%
  mutate( name = rownames(.) ) %>%
  arrange( membership, desc( degree ) ) %>%
  select( name ) %>%
  as.vector %>%
  inner_join( karate_df$vertices )
karate_df$edges$from <- match(karate_df$edges$from,karate_df$vertices$name) - 1
karate_df$edges$to <- match(karate_df$edges$to,karate_df$vertices$name) - 1

adjacencyNetwork(
  Links = karate_df$edges,
  Nodes = karate_df$vertices[,c(2,1)],
  Source = "from",
  Target = "to",
  NodeID = "name",
  Group = "Faction",
  margin = list(left=100, top=100),
  width = 500,
  height = 500
) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example05.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 650
      , height = 650
      , overflow = "hidden"
   ))
```

As you can see, I have a long way to go with this.  The dynamic sorting in the Mike Bostock example is probably first to implement.  I'd love help, comments, feedback, anything...

## Thanks

Thanks Mike Bostock for `d3` and another great example using it.  Thanks Elijah Meeks for all his `d3` posts, book, examples, and layouts.  Thanks Joe Cheng for `d3heatmap`.

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

