---
title: "Week41 - railroadR"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set( 
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## This Week's Widget - railroadR

[Railroad diagrams](https://en.wikipedia.org/wiki/Syntax_diagram) are generally used in computer science to represent language syntax or grammar.  However, I just know these can be used for other purposes. I just have not had the time for creativity or inspiration.  Please, please let me know if you have ideas.

Anyways, Tab Atkins, Jr. created this nice library [railroad-diagrams](https://github.com/tabatkins/railroad-diagrams/blob/master/example.html) to make these good looking diagrams.  I saw that it had a Python port, so I figured we needed it in `R` as an `htmlwidget`.

Even if you don't need `railroad-diagrams`, this little `htmlwidget` is a great example of the magic of non-standard evaluation, specifically with the Hadley Wickham package [`lazyeval`](https://github.com/hadley/lazyeval).  Using `lazyeval` allows a `railroadR` user to specify a diagram without a whole bunch of quotes and backslashes.  Have a look at these [lines](https://github.com/timelyportfolio/railroadR/blob/master/R/railroad.R#L14) if interested.

```{r eval = FALSE}
diagram <- lazyeval::auto_name(lazyeval::lazy_dots(...))

# forward options using x
x = list(
  diagram = unname(lapply(
    diagram,
    function(x) as.character(unclass(x)[1])
  ))
)
```

## Installation

`railroadR` is not on CRAN (happy to add if anybody requests), so please use `devtools::install_github` to install.

```{r eval = FALSE}
devtools::install_github("timelyportfolio/railroadR")
```

## Examples

Since I did not have time to explore creative uses of `railroadR`, I'll recreate the more complicated `railroad-diagram` [examples](http://www.xanthir.com/etc/railroad-diagrams/example.html).

### UNICODE-RANGE

```{r "example01", eval = F}
#devtools::install_github("timelyportfolio/railroadR")

library(railroadR)

#from https://github.com/tabatkins/railroad-diagrams/blob/master/example.html

railroad(
	Choice(0,
		'U',
		'u'),
	'+',
	Choice(0,
		Sequence(OneOrMore(NonTerminal('hex digit'), Comment('1-6 times'))),
		Sequence(
			ZeroOrMore(NonTerminal('hex digit'), Comment('1-5 times')),
			OneOrMore('?', Comment('1 to (6 - digits) times'))),
		Sequence(
			OneOrMore(NonTerminal('hex digit'), Comment('1-6 times')),
			'-',
			OneOrMore(NonTerminal('hex digit'), Comment('1-6 times'))))
)
```

```{r "example01_run", echo = F}
#devtools::install_github("timelyportfolio/railroadR")

library(railroadR)

#from https://github.com/tabatkins/railroad-diagrams/blob/master/example.html

railroad(
	Choice(0,
		'U',
		'u'),
	'+',
	Choice(0,
		Sequence(OneOrMore(NonTerminal('hex digit'), Comment('1-6 times'))),
		Sequence(
			ZeroOrMore(NonTerminal('hex digit'), Comment('1-5 times')),
			OneOrMore('?', Comment('1 to (6 - digits) times'))),
		Sequence(
			OneOrMore(NonTerminal('hex digit'), Comment('1-6 times')),
			'-',
			OneOrMore(NonTerminal('hex digit'), Comment('1-6 times'))))
,height = 200, width = "95%") %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = "90%"
      , height = 320
      , overflow = "hidden"
   ))
```

### SQL

```{r "example02", eval = F}
#devtools::install_github("timelyportfolio/railroadR")

library(railroadR)

#from https://github.com/tabatkins/railroad-diagrams/blob/master/example.html

railroad(
	Stack(
		Sequence(
			'SELECT',
			Optional('DISTINCT', 'skip'),
			Choice(0, 
				'*', 
				OneOrMore(
					Sequence(NonTerminal('expression'), Optional(Sequence('AS', NonTerminal('output_name')))), 
					','
				)
			),
			'FROM',
			OneOrMore(NonTerminal('from_item'), ','),
			Optional(Sequence('WHERE', NonTerminal('condition')))
		),
		Sequence(
			Optional(Sequence('GROUP BY', NonTerminal('expression'))),
			Optional(Sequence('HAVING', NonTerminal('condition'))),
			Optional(Sequence(
				Choice(0, 'UNION', 'INTERSECT', 'EXCEPT'),
				Optional('ALL'),
				NonTerminal('select')
			))
		),
		Sequence(
			Optional(Sequence(
				'ORDER BY',
				OneOrMore(Sequence(NonTerminal('expression'), Choice(0, Skip(), 'ASC', 'DESC')), ','))
			),
			Optional(Sequence(
				'LIMIT',
				Choice(0, NonTerminal('count'), 'ALL')
			)),
			Optional(Sequence('OFFSET', NonTerminal('start'), Optional('ROWS')))
		)
	)
)
```

```{r "example02_run", echo = F}
#devtools::install_github("timelyportfolio/railroadR")

library(railroadR)

#from https://github.com/tabatkins/railroad-diagrams/blob/master/example.html

railroad(
	Stack(
		Sequence(
			'SELECT',
			Optional('DISTINCT', 'skip'),
			Choice(0, 
				'*', 
				OneOrMore(
					Sequence(NonTerminal('expression'), Optional(Sequence('AS', NonTerminal('output_name')))), 
					','
				)
			),
			'FROM',
			OneOrMore(NonTerminal('from_item'), ','),
			Optional(Sequence('WHERE', NonTerminal('condition')))
		),
		Sequence(
			Optional(Sequence('GROUP BY', NonTerminal('expression'))),
			Optional(Sequence('HAVING', NonTerminal('condition'))),
			Optional(Sequence(
				Choice(0, 'UNION', 'INTERSECT', 'EXCEPT'),
				Optional('ALL'),
				NonTerminal('select')
			))
		),
		Sequence(
			Optional(Sequence(
				'ORDER BY',
				OneOrMore(Sequence(NonTerminal('expression'), Choice(0, Skip(), 'ASC', 'DESC')), ','))
			),
			Optional(Sequence(
				'LIMIT',
				Choice(0, NonTerminal('count'), 'ALL')
			)),
			Optional(Sequence('OFFSET', NonTerminal('start'), Optional('ROWS')))
		)
	)
,height = 600, width = "95%") %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example02.html"
      , libdir = "lib"
      , selfcontained=F
      , width = "90%"
      , height = 600
      , overflow = "hidden"
   ))
```


## Thanks

Thanks Tab Atkins, Jr. for the brilliant [railroad-diagrams](https://github.com/tabatkins/railroad-diagrams).

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

