---
title: "Week36 - stockchartR"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## This Week's Widget - stockchartR

I intentionally do not talk much finance here at Building Widgets, but for those who are unaware of [TimelyPortfolio](http://timelyportfolio.blogspot.com), finance pays the bills for me.  Although `d3.js` was released in 2010 (see this [first Github commit](https://github.com/mbostock/d3/commit/b2aa3c7355f46634738b46789c649c4e2da94e3e), we did not see robust finance-targeted `d3.js` visualization libraries until last year.  Now suddenly we have three really solid, very well-built options.

1. [techan.js](http://techanjs.org/)
2. [d3fc](http://d3fc.io/)
3. [react-stockcharts](http://rrag.github.io/react-stockcharts/)

I have wanted to `htmlwidget`-ize all of these but to do a good job on any of them would require at least a month of dedicated effort.  However, I just couldn't resist doing a sparse teaser implementation of `react-stockcharts`. I chose this one only because I have not done a [React](http://facebook.github.io/react/) `htmlwidget` yet.  I strongly encourage the interested reader to check out all three of the options listed above.  Also, please, please let me know if you want to collaborate on a `htmlwidget` for any of these.  All of them are great.

## Installation

```{r eval = FALSE}
devtools::install_github("timelyportfolio/stockchartR")
```

## Examples

As of now, this sparse wrapper provides a finance-based time scale, pan+zoom interaction, candlestick plotting, and a volume bar chart.  I'll say one more time just to be clear that `react-stockcharts` does much, much more.


```{r "example01", eval = FALSE}
# devtools::install_github("timelyportfolio/stockchartR")
library(stockchartR)
library(quantmod)

iwm <- getSymbols("IWM", auto.assign=FALSE, from="2014-01-01")
stockchart(adjustOHLC(iwm))
```

```{r "example01_run", echo = FALSE}
# devtools::install_github("timelyportfolio/stockchartR")
library(stockchartR)
library(quantmod)

iwm <- getSymbols("IWM", auto.assign=FALSE, from="2014-01-01")
stockchart(adjustOHLC(iwm),width = 620,height = 400)%>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 750
      , height = 480
      , overflow = "hidden"
   ))
```

<span style="font-size:.6em;font-style:italic;">source: Yahoo! Finance</span>


## Thanks

[Ragu Ramaswamy](https://github.com/rrag), thanks for your really fine work on [`react-stockcharts`](http://rrag.github.io/react-stockcharts/).  I can't wait to implement and take advantage of all that you have built into this amazing library.

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

