---
title: "Week15 - List Viewer"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## htmlwidgets News This Week

`htmlwidgets` now uses [`jsonlite`](https://github.com/jeroenooms/jsonlite).  I recommend anyone serious about their widgets to read through the discussion of [pull #28](https://github.com/ramnathv/htmlwidgets/pull/28).  Whether you care about the details or not, I would encourage updatating `jsonlite`, `htmlwidgets`, and `Shiny` with

```{r}

```

For all the newest and updated `htmlwidgets`, just do this [Github search](https://github.com/search?o=desc&q=htmlwidgets&s=indexed&type=Code), and you'll quickly get up to date.

## This Week's Widget - `listviewer`

Last week's [`exportwidget`](http://www.buildingwidgets.com/blog/2015/4/9/week-14-exporting-widget) got almost no love, so I'm hoping this one finds a loving home.  I have not been able to stop using it, since the first hour of creation, so it has quickly become a key component in my workflow.

Nested data gets messy, inscrutable, and difficult to follow very quickly.  `lists` and `environments` in R are the primary source of this nested data.  Often, an interactive visual look provides the insight and understanding one needs.  `listviewer` gives us this view with just one line of code.  We can use it to inspect `htmlwidgets` or `lattice` plots or `par()` settings or your `.GlobalEnv`.

Thanks so much to the [Jos de Jong](http://josdejong.com/), the author of [`jsoneditor`](https://github.com/josdejong/jsoneditor) on which `listviewer` relies for all its awesomeness.  If you look at the source of `listviewer`, you will notice how little code I had to write to wrap this as an `htmlwidget`.


## Quick Installation

As with almost all widgets posted here, `listviewer` is not on CRAN, so for now please install with `devtools::install_github`.  Given enough interest, I'm happy to put in the effort to make this or any others CRAN-worthy.  Just let me know.

```{r eval = F}
devtools::install_github("timelyportfolio/listviewer")
```

## Examples

### par() what?

When I first learned `R` I struggled with all the stuff in `par()`.  Let's see if this makes a little more sense.

```{r "example01", echo=T, eval = F}
library(listviewer)

jsonedit( par() )
```

```{r "example01_run", echo=F, eval = T}
library(listviewer)

jsonedit( par(), height = 400, width = 400 ) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 520
      , height = 520
      , overflow = "hidden"
    )
  )
```

### Example with another widget `streamgraph`



```{r "example2", echo = T, eval = F}

```
```{r "example2_run", echo = F, eval = T}

```

### Example with multiple widgets


```{r "example03", echo = T, eval = F}

```
```{r "example03_run", echo = F, eval = T}

```

### Example | Headless with `webshot`


```{r "example04", echo = T, eval = F}

```

```{r "example04_run", echo = F, eval = T}

```

## Thanks

Thanks so much for all of this

- [`jsoneditor`](https://github.com/josdejong/jsoneditor) from Jos de Jong
- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

