---
title: "Week15 - List Viewer"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## htmlwidgets News This Week

`htmlwidgets` now uses [`jsonlite`](https://github.com/jeroenooms/jsonlite).  I recommend anyone serious about their widgets to read through the discussion of [pull #28](https://github.com/ramnathv/htmlwidgets/pull/28).  Whether you care about the details or not, I would encourage updatating `jsonlite`, `htmlwidgets`, and `Shiny` with

```{r eval=F, echo=T}
devtools::install_github(c(
  'jeroenooms/jsonlite'
  , 'rstudio/shiny'
  , 'ramnathv/htmlwidgets'
))
```

For all the newest and updated `htmlwidgets`, just do this [Github search](https://github.com/search?o=desc&q=htmlwidgets&s=indexed&type=Code), and you'll quickly get up to date.

## This Week's Widget - `listviewer`

Last week's [`exportwidget`](http://www.buildingwidgets.com/blog/2015/4/9/week-14-exporting-widget) got almost no love, so I'm hoping this one finds a loving home.  I have not been able to stop using it, since the first hour of creation, so it has quickly become a key component in my workflow.

With a [tweet](https://twitter.com/JennyBryan/status/586327100769968128), Jennifer Bryan inspired this widget.  Nested data gets messy, inscrutable, and difficult to follow very quickly.  `lists` and `environments` in R are the primary source of this nested data.  Often, an interactive visual look provides the insight and understanding one needs.  `listviewer` gives us this view with just one line of code.  We can use it to inspect `htmlwidgets` or `lattice` plots or `par()` settings or your `.GlobalEnv`.

Thanks so much to the [Jos de Jong](http://josdejong.com/), the author of [`jsoneditor`](https://github.com/josdejong/jsoneditor) on which `listviewer` relies for all its awesomeness.  If you look at the source of `listviewer`, you will notice how little code I had to write to wrap this as an `htmlwidget`.


## Quick Installation

As with almost all widgets posted here, `listviewer` is not on CRAN, so for now please install with `devtools::install_github`.  Given enough interest, I'm happy to put in the effort to make this or any others CRAN-worthy.  Just let me know.

```{r eval = F}
devtools::install_github("timelyportfolio/listviewer")
```

## Examples

### What's my `par`?

When I first learned `R` I struggled with all the stuff in `par()`.  Let's see if this makes a little more sense.

```{r "example01", echo=T, eval = F}
library(listviewer)

jsonedit( par() )
```

```{r "example01_run", echo=F, eval = T}
library(listviewer)

jsonedit( par(), height = 400, width = 400 ) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 520
      , height = 520
      , overflow = "hidden"
    )
  )
```

### What's in my `lattice`?

`lattice` plots become `trellis` objects.  Let's peek inside.  If you're thinking let's do the same with `ggplot2`, you'll need to write a little more code to handle `S3`, `S4`, and `environments`.

```{r "example2", echo = T, eval = F}
library(lattice)
library(listviewer)

jsonedit(
  xyplot( y~x, data.frame( x = 1:10, y = 1:10 ), type = "b" )
)
```
```{r "example2_run", echo = F, eval = T}
library(lattice)
library(listviewer)

jsonedit(
  xyplot( y~x, data.frame( x = 1:10, y = 1:10 ), type = "b" )
  ,height = 600
  ,width = 400
)  %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example02.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 520
      , height = 720
      , overflow = "hidden"
    )
  )
```

### What's in my `broom`?

David Robinson's [`broom` package](http://varianceexplained.org/r/broom-slides/) tidies up our messy data in `R`.  Let's see the difference with `jsonedit`.


```{r "example03", echo = T, eval = F}
jsonedit(
  list(
    "messy" = lm(mpg~factor(cyl),mtcars)
    ,"broom" = list(
      "tidy" = broom::tidy(lm(mpg~factor(cyl),mtcars))
      ,"augment" = broom::augment(lm(mpg~factor(cyl),mtcars))
    )
  )
)
```
```{r "example03_run", echo = F, eval = T}
jsonedit(
  list(
    "messy" = lm(mpg~factor(cyl),mtcars)
    ,"broom" = list(
      "tidy" = broom::tidy(lm(mpg~factor(cyl),mtcars))
      ,"augment" = broom::augment(lm(mpg~factor(cyl),mtcars))
    )
  ), height = 600, width = 400
) %>>%
    (
    htmlwidgets::as.iframe(
      .
      ,file="example03.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 520
      , height = 720
      , overflow = "hidden"
    )
  )
```

### What's in my `htmlwidget`?

I hope you can see some of the power of this `htmlwidget`.  As a final example, let's go meta and `jsonedit` our `jsonedit`.


```{r "example04", echo = T, eval = F}
library(listviewer)

jsonedit(jsonedit())
```

```{r "example04_run", echo = F, eval = T}
library(listviewer)

jsonedit(jsonedit(),height=400,width = 400)  %>>%
    (
    htmlwidgets::as.iframe(
      .
      ,file="example04.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 520
      , height = 520
      , overflow = "hidden"
    )
  )
```

### Shiny?

`listviewer` sort of works with `Shiny`.  `modify` and `save` for `listviewer` is the next frontier.  See the [`Readme.md` ](https://github.com/timelyportfolio/listviewer/blob/master/README.md).

## Thanks

Thanks so much

- [`jsoneditor`](https://github.com/josdejong/jsoneditor) from Jos de Jong
- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

