---
title: "Week28 - d3treeR"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## htmlwidgets News This Week

A <span style="font-size:120%">gallery is really close</span>.  Just a little longer until its release, for all the newest and updated `htmlwidgets`, continue to search [Github search](https://github.com/search?o=desc&q=htmlwidgets&s=indexed&type=Code).

Also, I should note that [Week 2's](www.buildingwidgets.com/blog/2015/1/15/week-02-) widget [svgPanZoom](https://github.com/timelyportfolio/svgPanZoom) achieved [CRAN](https://cran.r-project.org/web/packages/svgPanZoom).  Just ask if you would like me to work through the CRAN process for any others.

## This Week's Widget - `d3treeR`

[Martijn Tennekes](https://github.com/mtennekes) is one of my favorite `R` package authors.  At [Vis2014](http://ieeevis.org/year/2014/info/overview-amp-topics/paper-sessions), he presented his work with Edwin de Jonge on [treemaps](https://github.com/mtennekes/treemap) and [hierarchical colors](https://github.com/mtennekes/hierCols) both of which have `R` packages.

[Bill White](http://www.billdwhite.com/wordpress/) has done some very interesting work with `d3.js`.  He has attacked the treemap problem on the JavaScript side and explained it all very well with source code (generously [MIT licensed](http://www.billdwhite.com/wordpress/source-code-license/)) in his post [D3 TREEMAP WITH TITLE HEADERS](http://www.billdwhite.com/wordpress/2012/12/16/d3-treemap-with-title-headers/).

Tennekes' treemap in `R` gives us a powerful set of tools to transform, analyze, and color treemaps.  Let's combine all this functionality with the nice interactivity of Bill White's `d3.js` treemaps in this week's `htmlwidget` [d3treeR](https://github.com/timelyportfolio/d3treeR).

## Quick Installation

`d3treeR` is not yet on CRAN and won't be unless someone asks, so for now please install with `devtools::install_github`. 

```{r eval = F}
devtools::install_github("timelyportfolio/d3treeR")
```


## Examples


### `treemap` as data

As stated in the introduction, `d3treeR` is built to work seamlessly with `treemaps` from the `R` treemap package.

```{r "example01", eval = FALSE}
# devtools::install_github("timelyportfolio/d3treeR")
library(treemap)
library(d3treeR)

# first example from ?treemap

data(GNI2010)
d3tree(
  treemap(
     GNI2010,
     index=c("continent", "iso3"),
     vSize="population",
     vColor="GNI",
     type="value"
   )
  ,rootname = "World"
)
```

```{r "example01_run", eval = TRUE, echo = FALSE}
# devtools::install_github("timelyportfolio/d3treeR")
library(treemap)
library(d3treeR)

# first example from ?treemap

data(GNI2010)
d3tree(
  treemap(
     GNI2010,
     index=c("continent", "iso3"),
     vSize="population",
     vColor="GNI",
     type="value"
   )
  ,rootname = "World"
  ,height = 520
  ,width = 620
) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 800
      , height = 670
      , overflow = "hidden"
   ))
```

```{r "example02", eval = FALSE}
# devtools::install_github("timelyportfolio/d3treeR")
library(treemap)
library(d3treeR)

# last example from ?treemap

data(business)
business$employees.growth <- business$employees - business$employees.prev
d3tree(
  # Brewer's Red-White-Grey palette reversed with predefined range
  treemap(business,
          index=c("NACE1", "NACE2"),
          vSize="employees",
          vColor="employees.growth",
          type="value",
          palette="-RdGy",
          range=c(-30000,30000))
  ,rootname = "World"
)
```

```{r "example02_run", eval = TRUE, echo = FALSE}
library(treemap)
library(d3treeR)

# last example from ?treemap

data(business)
business$employees.growth <- business$employees - business$employees.prev
d3tree(
  # Brewer's Red-White-Grey palette reversed with predefined range
  treemap(business,
          index=c("NACE1", "NACE2"),
          vSize="employees",
          vColor="employees.growth",
          type="value",
          palette="-RdGy",
          range=c(-30000,30000))
  ,rootname = "World"
  ,height = 520
  ,width = 620
) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example02.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 800
      , height = 670
      , overflow = "hidden"
   ))
```

### `JSON` as data

We can use `JSON` or lists from `R` to provide our hierarchy.  Let's use the ubiquitous `flare.json`.

```{r "example03", eval = FALSE}
library(treemap)
library(d3treeR)

  # with flare.json from Mike Bostock

d3tree(
  "http://bl.ocks.org/mbostock/raw/4063582/raw/flare.json"
  ,id = "name"
  ,height = 520
  ,width = 620
)
```

```{r "example03_run", eval = TRUE, echo = FALSE}
library(treemap)
library(d3treeR)

  # with flare.json from Mike Bostock

d3tree(
  "http://bl.ocks.org/mbostock/raw/4063582/raw/flare.json"
  ,id = "name"
  ,height = 520
  ,width = 620
) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example03.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 800
      , height = 670
      , overflow = "hidden"
   ))
```

## Thanks

Thanks so much to Martijn Tennekes, Edwin de Jonge, Mike Bostock, and Bill White.

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

