---
title: "Week38 - datacomb"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set( 
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## This Week's Widget - datacomb

I was blown away when an August 21 \@ChrisPolis [tweet](https://twitter.com/ChrisPolis/status/634604023044182016) led me to this animated GIF.

<img alt="animated gif of datacomb" src = "https://rawgit.com/cmpolis/datacomb/master/demo/poc-demo.gif" width = "750px">

Then, I was even more delighted when I saw the magic word `htmlwidget` on the roadmap/todos.  Just a month later thanks to magic/diligence of Chris Polis we can make a `datacomb` interactive table viewer in one-line of `R`.

<span style="font-size:120%;font-style:italic;font-weight:bold;">Note, you will notice this is not done and there is much left to do</span>, but I just couldn't wait to show this one off.

## Installation

```{r eval = FALSE}
devtools::install_github("cmpolis/datacomb",subdir="pkg")
```

## Example

In one line, we get a rich quick interactive overview of pretty-"big" data.

```{r "example01", eval = FALSE}
# devtools::install_github("cmpolis/datacomb", subdir="pkg")
library(datacomb)
data(diamonds, package="ggplot2")
Datacomb(diamonds)
```

```{r "example01_run", echo = FALSE}
library(datacomb)
data(diamonds, package="ggplot2")
Datacomb(diamonds,height=520,width=600)%>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 750
      , height = 650
      , overflow = "hidden"
   ))
```


## Thanks

`datacomb` on both the `JS` and `R` side is all the work of [Chris Polis \@cmpolis](https://github.com/cmpolis), so please direct all thanks, praise, and attribution to him.  `datacomb` has amazed me and many others since we first saw the animated GIF preview.  Thanks Chris!

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

