---
title: "Week52 - d3kit_timeline"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set( 
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```


The clock is ticking rapidly--so rapidly that I won't be able to do the full year summary, meta-filled post that I had hoped in this final week of the htmlwidget/week project started with this January 2, 2015 post [Can I Commmit?](http://www.buildingwidgets.com/blog/2015/1/2/can-i-commit).  I'll stick with the traditional post format for the release of week 52's widget `d3kit_timeline`.  Then in 2016, I'll try to complete the post I had planned.  Thanks so much for all those who have followed along.

## This Week's Widget - d3kit_timeline

At this point in the year, we often focus on time, so I thought this week's `htmlwidget` should also focus on time.  Ramnath Vaidyanathan (one of the primary htmlwidgets authors) suggested [`TimelineJS`](https://timeline.knightlab.com/) which he had previously [wrapped in `rCharts`](http://ramnathv.github.io/sachin100s/) and I used for this [history of `R` releases](http://timelyportfolio.github.io/rCharts_timeline_r/).  However, I got distracted by [Krist Wongsuphasawat's / \@kristw ](https://twitter.com/kristw) `d3.js` based timeline [`d3kit_timeline`](https://github.com/kristw/d3kit-timeline) that uses his highly starred, newly released [labella.js](http://twitter.github.io/labella.js/index.html).  It is not quite as flashy as `TimelineJS`, but I think it is highly useful.

I try to combine similar `htmlwidgets` into packages when possible.  `d3kit_timeline` seemed to fit nicely into [`timelineR`](https://github.com/timelyportfolio/timelineR) that also contains [week 39's `timeline`](http://www.buildingwidgets.com/blog/2015/10/2/week-39-timeliner).

## Installation

This is not on CRAN, so to install we will need some help from `devtools::install_github`.  **We will also require the newest `htmlwidgets`** to use the new instance-bound syntax.

```{r eval = FALSE}
devtools::install_github("ramnathv/htmlwidgets")
devtools::install_github("timelyportfolio/timelineR")
```

## Examples

### Simple Example

Let's start by replicating one of the [examples](http://kristw.github.io/d3kit-timeline/) from `d3kit-timeline`.  For those who want more, I have [replicated all of these examples](https://github.com/timelyportfolio/timelineR/blob/master/inst/examples/example_d3kit_timeline.R) in `?d3kit_timeline`.

I really wanted to focus this week on being very flexible with the arguments to bridge the `R` and JavaScript worlds.  To accomplish this, I spent a lot of time on this [ugly bit of code](https://github.com/timelyportfolio/timelineR/blob/master/R/d3kit_timeline.R#L87-L168) inspired by these [lines](https://github.com/jcheng5/d3scatter/blob/master/R/d3scatter.R#L10-L16) from Joe Cheng at RStudio in his test `htmlwidget` `d3scatter`.

```{r "example01", eval = F}
# devtools::install_github("ramnathv/htmlwidgets")
# devtools::install_github("timelyportfolio/timelineR")

library(timelineR)

# replicate example from http://kristw.github.io/d3kit-timeline/#
# define starwars release data used in all the examples
starwars_data <- data.frame(
  time = c(
    "1977-04-25",
    "1980-04-17",
    "1984-04-25",
    "1999-04-19",
    "2002-04-16",
    "2005-04-19",
    "2015-11-18"
  ),
  episode = c(4,5,6,1,2,3,7),
  name = c(
    'A New Hope',
    'The Empire Strikes Back',
    'Return of the Jedi',
    'The Phantom Menace',
    'Attack of the Clones',
    'Revenge of the Sith',
    'The Force Awakens'
  ),
  stringsAsFactors = FALSE
)

d3kit_timeline(
  starwars_data,
  direction = "right",
  # time is default but show as example of flexible argument types
  timeFn = ~time,
  textFn = htmlwidgets::JS(
"
function(d){
    return new Date(d.time).getFullYear() + ' - ' + d.name;
}
"
  ),
  width = 400,
  height = 250
)
```

```{r "example01_run", echo = F}
# devtools::install_github("ramnathv/htmlwidgets")
# devtools::install_github("timelyportfolio/timelineR")

library(timelineR)

# replicate example from http://kristw.github.io/d3kit-timeline/#
# define starwars release data used in all the examples
starwars_data <- data.frame(
  time = c(
    "1977-04-25",
    "1980-04-17",
    "1984-04-25",
    "1999-04-19",
    "2002-04-16",
    "2005-04-19",
    "2015-11-18"
  ),
  episode = c(4,5,6,1,2,3,7),
  name = c(
    'A New Hope',
    'The Empire Strikes Back',
    'Return of the Jedi',
    'The Phantom Menace',
    'Attack of the Clones',
    'Revenge of the Sith',
    'The Force Awakens'
  ),
  stringsAsFactors = FALSE
)

d3kit_timeline(
  starwars_data,
  direction = "right",
  # time is default but show as example of flexible argument types
  timeFn = ~time,
  textFn = htmlwidgets::JS(
"
function(d){
    return new Date(d.time).getFullYear() + ' - ' + d.name;
}
"
  ),
  width = 400,
  height = 250
) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 570
      , height = 420
      , overflow = "hidden"
    )
  )
```

### Use With `xts`

I love `xts`, and `xts` is one of the reasons I finally learned `R`, so I wanted `d3kit_timeline` to play nicely with these time series objects.

Off topic from `R`, AJ McCarron finally broke the 1987 winless streak for starting NFL quarterbacks from Alabama ([see article](http://ftw.usatoday.com/2015/12/aj-mccarron-breaks-alabama-quarterback-win-nfl-drought-which-college-has-next-longest-qb-drought)).  I thought it would be interesting to see the Alabama quarterbacks that played (all won) in the Superbowl with our new `d3kit_timeline`.  I combined data into an `xts` object from this [NFL.com article](http://www.nfl.com/news/story/0ap2000000321044/article/notre-dame-stanford-have-most-super-bowl-starts-at-qb) and this [wikipedia data](https://en.wikipedia.org/wiki/List_of_Super_Bowl_champions).

```{r "example02", eval = F}
# devtools::install_github("ramnathv/htmlwidgets")
# devtools::install_github("timelyportfolio/timelineR")

library(timelineR)
library(xts)
library(pipeR)

# sources:
# http://www.nfl.com/news/story/0ap2000000321044/article/notre-dame-stanford-have-most-super-bowl-starts-at-qb
# https://en.wikipedia.org/wiki/List_of_Super_Bowl_champions

alqb_xts <- as.xts(
  data.frame(
    quarterback = c(rep("Bart Starr",2),"Joe Namath", "Kenny Stabler"),
    team = c(rep("Packers",2),"Jets","Raiders"),
    color = c(rep("#FFB612",2),"#0C371D","#000"),
    stringsAsFactors = FALSE
  ),
  order.by = as.Date(c(
    "1967-01-15",
    "1968-01-14",
    "1969-01-12",
    "1977-01-09"
  ))
)

colorJS <- htmlwidgets::JS("function(d){return d.color;}")

d3kit_timeline(
  alqb_xts,
  direction = "down",
  textFn = htmlwidgets::JS(
"
function(d){
    return d.quarterback + ' - ' + d.team;
}
"
  ),
  # color probably needs to be treated like the *Fn arguments
  #  for ultimate flexibility
  dotColor = colorJS,
  linkColor = colorJS,
  labelTextColor = "#FFF",
  labelBgColor = colorJS,
  dotRadius = 5,
  labella = list(maxPos = 600),
  margin = list(left = 20, right = 50, top = 20, bottom = 40),
  width = 500,
  height = 250
) %>>%
  add_axis( ticks = 7  )
```

```{r "example02_run", echo = F}
# devtools::install_github("ramnathv/htmlwidgets")
# devtools::install_github("timelyportfolio/timelineR")

library(timelineR)
library(xts)
library(pipeR)

# sources:
# http://www.nfl.com/news/story/0ap2000000321044/article/notre-dame-stanford-have-most-super-bowl-starts-at-qb
# https://en.wikipedia.org/wiki/List_of_Super_Bowl_champions

alqb_xts <- as.xts(
  data.frame(
    quarterback = c(rep("Bart Starr",2),"Joe Namath", "Kenny Stabler"),
    team = c(rep("Packers",2),"Jets","Raiders"),
    color = c(rep("#FFB612",2),"#0C371D","#000"),
    stringsAsFactors = FALSE
  ),
  order.by = as.Date(c(
    "1967-01-15",
    "1968-01-14",
    "1969-01-12",
    "1977-01-09"
  ))
)

colorJS <- htmlwidgets::JS("function(d){return d.color;}")

d3kit_timeline(
  alqb_xts,
  direction = "down",
  textFn = htmlwidgets::JS(
"
function(d){
    return d.quarterback + ' - ' + d.team;
}
"
  ),
  # color probably needs to be treated like the *Fn arguments
  #  for ultimate flexibility
  dotColor = colorJS,
  linkColor = colorJS,
  labelTextColor = "#FFF",
  labelBgColor = colorJS,
  dotRadius = 5,
  labella = list(maxPos = 600),
  margin = list(left = 20, right = 50, top = 20, bottom = 40),
  width = 500,
  height = 250
) %>>%
  add_axis( ticks = 7  ) %>>%
  (
    htmlwidgets::as.iframe(
      .
      ,file="example02.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 620
      , height = 420
      , overflow = "hidden"
    )
  )
```

## Thanks

Thanks [Krist Wongsuphasawat](https://twitter.com/kristw) for his beautiful work on Labella.js and d3kit-timeline.

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

