---
title: "Week06 - bokeh now for R"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
---

```{r "setup",echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
  #, results = 'asis'
)

library(pipeR)
```

## htmlwidgets News This Week
-[\@hrbrmstr](https://github.com/hrbrmstr) - d3 streamgraph layout [`streamgraph`](https://github.com/hrbrmstr/streamgraph).  Bob Rudis builds on his list of authored widgets with this very nice `htmlwidget` for streamgraphs.  See the [announcement](http://hrbrmstr.github.io/streamgraph/) for a through discussion.

-[\@smartinsightsfromdata](https://github.com/smartinsightsfromdata) - [`rpivotTable`](https://github.com/smartinsightsfromdata/rpivotTable) provides a `htmlwidget` for the very well done [PivotTable.js](http://nicolas.kruchten.com/pivottable/examples/) by [Nicholas Krutchen](http://nicolas.kruchten.com/).  Not only will Excel users feel more at home now, but also there are lots of very interesting views, such as treemaps and heatmaps.  If you still like old-fashioned `tables` and their derivatives in R like I do, please let me know, and maybe we can convince the author to add the couple lines to support these very handy and far-from-obsolete beauties.


## This Week's Widget - `rbokeh` | bokeh for R

Python and even Scala and Julia have [`Bokeh`](http://bokeh.pydata.org/en/latest/), so R should too.  Fortunately, [Ryan Hafen](https://twitter.com/hafenstats) from Purdue and [Tessara](http://tessera.io/) has been applying his very thorough R skills to make [`rbokeh`](http://hafen.github.io/rbokeh), a professionally built `htmlwidget` for `Bokeh`.  Just look at that site and the documentation, and I think you'll quickly get as excited as I am for the potential of `rbokeh`.

I should **note that Ryan deserves all the credit for `rbokeh`**.  I simply chipped in with some additional examples, testing, lots of questions, and minor pull requests.  I hope I can also help implement a bar chart and full seamless shiny integration.  Feel free to contribute any way you can.  Every little bit helps.

## Examples

I understand how hard it is to motivate R users spoiled by `ggplot2`, `lattice`, and even base `graphics`.  Borrowing from these great libraries, Ryan has integrated many good bits into `rbokeh`.  Enough writing, let's see some examples.  You'll notice `rbokeh` loves modern R pipes.

```{r "example01", eval = F}
# devtools::install("hafen/rbokeh")

library(rbokeh)
# basic pipes built in but I love pipeR
library(pipeR)

# diamonds is always my goto for kind of big data
data(diamonds, package="ggplot2")

figure( width = 600, height = 600, title = "Diamonds from ggplot2" ) %>>%
  ly_hexbin(
    carat
    ,log(price)
    ,data = diamonds
  )
```


```{r "example01_run", echo = F}
# devtools::install("hafen/rbokeh")

library(rbokeh)
# basic pipes built in but I love pipeR
library(pipeR)

# diamonds is always my goto for kind of big data
data(diamonds, package="ggplot2")

figure( width = 600, height = 600, title = "Diamonds from ggplot2" ) %>>%
  ly_hexbin(
    carat
    ,log(price)
    ,data = diamonds
  ) %>>%
  (
   htmlwidgets::as.iframe(
      rbokeh:::plot.BokehFigure(.)
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 750
      , height = 750
      , overflow = "hidden"
   ))
```

`Bokeh` renders to canvas, so it can handle lots of points.

```{r "example02", eval = F}
figure( width = 600, height = 400, title = "Diamonds from ggplot2" ) %>>%
  ly_points(
    x = carat
    , y = price
    , data = diamonds
    , color = color
    , size = 2
  )
```

```{r "example02_run", echo = F}
figure( width = 600, height = 400, title = "Diamonds from ggplot2" ) %>>%
  ly_points(
    x = carat
    , y = price
    , data = diamonds
    , color = color
    , size = 2
  ) %>>%
  (
   htmlwidgets::as.iframe(
      rbokeh:::plot.BokehFigure(.)
      ,file="example02.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 750
      , height = 550
      , overflow = "hidden"
   ))
```

`lm` plugs right into `rbokeh`.

```{r "example03", eval = F}
figure( width = 700, height = 400, title = "Diamonds lm with rbokeh" ) %>>%
  ly_points(
    x = carat
    , y = price
    , data = diamonds
    , color = color
    , size = 2
  ) %>>%
  ly_abline(
    lm( price ~ carat, diamonds )
    ,type = 2
    ,width = 3
  )
```
```{r "example03_run", echo = F}
figure( width = 600, height = 400, title = "Diamonds lm with rbokeh" ) %>>%
  ly_points(
    x = carat
    , y = price
    , data = diamonds
    , color = color
    , size = 2
  ) %>>%
  ly_abline(
    lm( price ~ carat, diamonds )
    ,type = 2
    ,width = 3
  ) %>>%
  (
   htmlwidgets::as.iframe(
      rbokeh:::plot.BokehFigure(.)
      ,file="example03.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 750
      , height = 550
      , overflow = "hidden"
   ))
```

Now would be a great time to show `facets` with `rbokeh`.  `grid_plot` gives us a way to achieve these, but `facets` require a little extra effort.  But hey, we all need to know `*apply` anyways.  Let's split the plot above by clarity.

```{r "example04", eval = F}
lapply(
  levels(diamonds$clarity)
  ,function(clar){
    figure( width = 300, height = 300, title = paste0("Clarity: ",clar )) %>>%
      ly_points(
        x = carat
        , y = price
        , data = subset(diamonds,clarity==clar)
        , color = color
        , size = 2
      ) %>>%
      ly_abline(
        lm( price ~ carat, subset(diamonds,clarity==clar) )
        ,type = 2
        ,width = 3
      )    
  }
) %>>%
  ( grid_plot( ., nrow = 2, ncol = 4, same_axes = T ) )
```
```{r "example04_run", echo = F}
lapply(
  levels(diamonds$clarity)
  ,function(clar){
    figure( width = 300, height = 300, title = paste0("Clarity: ",clar )) %>>%
      ly_points(
        x = carat
        , y = price
        , data = subset(diamonds,clarity==clar)
        , color = color
        , size = 2
      ) %>>%
      ly_abline(
        lm( price ~ carat, subset(diamonds,clarity==clar) )
        ,type = 2
        ,width = 3
      )    
  }
) %>>%
  ( grid_plot( ., nrow = 2, ncol = 4, same_axes = T ) ) %>>%
  (
   htmlwidgets::as.iframe(
      rbokeh:::plot.BokehGridPlot(.)
      ,file="example04.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 750
      , height = 750
      , overflow = "hidden"
   ))
```

### Examples from `ggplot2` & `lattice` books
`rbokeh` can do so much that I could write an entire book.  I have started by replicating many of the examples from `ggplot2` and `lattice` books in this [Gist](https://gist.github.com/timelyportfolio/8300c43ac43d772600df).


## Thanks

Thanks so much for all the work by

- [Ryan Hafen](https://github.com/hafen) for `rbokeh`
- team at [Continuum Analytics](http://continuum.io/)
- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
