---
title: "Week06 - bokeh now for R"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
---

```{r "setup",echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
  #, results = 'asis'
)

library(pipeR)
```

## htmlwidgets News This Week
-[\@hrbrmstr](https://github.com/hrbrmstr) - d3 streamgraph layout [`streamgraph`](https://github.com/hrbrmstr/streamgraph).  Bob Rudis builds on his list of authored widgets with this very nice `htmlwidget` for streamgraphs.  See the [announcement](http://hrbrmstr.github.io/streamgraph/) for a through discussion.

-[\@smartinsightsfromdata](https://github.com/smartinsightsfromdata) - [`rpivotTable`](https://github.com/smartinsightsfromdata/rpivotTable) provides a `htmlwidget` for the very well done [PivotTable.js](http://nicolas.kruchten.com/pivottable/examples/) by [Nicholas Krutchen](http://nicolas.kruchten.com/).  Not only will Excel users feel more at home now, but also there are lots of very interesting views, such as treemaps and heatmaps.  If you still like old-fashioned `tables` and their derivatives in R like I do, please let me know, and maybe we can convince the author to add the couple lines to support these very handy and far-from-obsolete beauties.


## This Week's Widget - `rbokeh` | bokeh for R

Python and even Scala and Julia have [`Bokeh`](http://bokeh.pydata.org/en/latest/), so R should too.  Fortunately, [Ryan Hafen](https://twitter.com/hafenstats) from Purdue and [Tessara](http://tessera.io/) has been applying his very thorough R skills to make [`rbokeh`](http://hafen.github.io/rbokeh), a professionally built `htmlwidget` for `Bokeh`.  Just look at that site and the documentation, and I think you'll quickly get as excited as I am for the potential of `rbokeh`.

I should **note that Ryan deserves all the credit for `rbokeh`**.  I simply chipped in with some additional examples, testing, lots of questions, and minor pull requests.  I hope I can also help implement a bar chart and full seamless shiny integration.  Feel free to contribute any way you can.  Every little bit helps.

## Examples

I understand how hard it is to motivate R users spoiled by `ggplot2`, `lattice`, and even base `graphics`.  Borrowing from these great libraries, Ryan has integrated many good bits into `rbokeh`.  Enough writing, let's see some examples.  You'll notice `rbokeh` loves modern R pipes.

```{r "example_01", eval = F}
# devtools::install("hafen/rbokeh")

library(rbokeh)


```



## Thanks

Thanks so much for all the work by

- [Ryan Hafen](https://github.com/hafen) for `rbokeh`
- team at [Continuum Analytics](http://continuum.io/)
- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
