---
title: "Week40 - networkD3 0.2.1 -> 0.2.4"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set( 
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## This Week's Widget - networkD3

[`networkD3`](http://christophergandrud.github.io/networkD3/) from [Christopher Gandrud](https://twitter.com/ChrisGandrud) is the second most popular `htmlwidget` based on Github stars, so I think it is important that it stays in *elite* shape.  Last week `networkD3` [hit CRAN](https://cran.r-project.org/web/packages/networkD3/index.html) with version `0.2.4`, so I'll claim `networkD3` as the `htmlwidget` of the week.  The main improvements are a cycle-supporting `sankeyNetwork` (no more infinite loops) and I hope a much better fitting `diagonalNetwork` and `radialNetwork`.

For those of you who don't know or forgot, I am using [Github issues on the buildingwidgets](https://github.com/timelyportfolio/buildingwidgets/issues) repo to list ideas for future `htmlwidgets`.  We only have 12 more in the year long adventure,  so I would love to hear from you if you have any particular widgets that intrigue you.

I also wanted to highlight [`scatterD3`](https://github.com/juba/scatterD3) which has turned into a very nice and powerful `htmlwidget`.

## Installation

`networkD3` is on CRAN, so installation is simple.

```{r eval = FALSE}
install.packages("networkD3")
```

For the latest and greatest, you can use `devtools::install_github`.

```{r eval = FALSE}
devtools::install_github("christophergandrud/networkD3")
```

## Examples

### diagonal/radialNetwork

`diagonalNetwork` and `radialNetwork` needed a little help automatically sizing themselves, so the user didn't need to endlessly fiddle with margins.  Also, the `margins` argument only supported a right margin.  Now, `margins` can be a single number if you only want right, but now `margins` also can be a `list` with any of `top`, `right`, `bottom`, or `left`.

Let's take a look at what happens if we make a tiny plot of a big network.  Change the size of your screen and see that we also get a little bit of "responsive"-ness.

```{r "example01", eval = FALSE}
library(networkD3)

URL <- paste0("https://cdn.rawgit.com/christophergandrud/networkD3/",
              "master/JSONdata//flare.json")

## Convert to list format
Flare <- jsonlite::fromJSON(URL, simplifyDataFrame = FALSE)

## Recreate Bostock example from http://bl.ocks.org/mbostock/4063550
diagonalNetwork(
  List = Flare, fontSize = 10, opacity = 0.9,
  height = "100%", width = "60%"
)
```

```{r "example01_run", eval = TRUE, echo = FALSE, fig.keep='none'}
library(networkD3)

URL <- paste0("https://cdn.rawgit.com/christophergandrud/networkD3/",
              "master/JSONdata//flare.json")

## Convert to list format
Flare <- jsonlite::fromJSON(URL, simplifyDataFrame = FALSE)

## Recreate Bostock example from http://bl.ocks.org/mbostock/4063550
diagonalNetwork(
  List = Flare, fontSize = 10, opacity = 0.9,
  height = "100%", width = "60%"
) %>>%  (
   htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = "90%"
      , height = 500
      , overflow = "hidden"
   ))
```



### sankeyNetwork

`sankeyNetwork` now needs a lot fewer arguments to work.  The function tries to be "smart" about guessing the structure of your data, such as in `Links`, it assumes `Source` will be column 1 and `Target` will be column 2 if not explicitly told where to look.

`sankeyNetwork` would result in a blank screen caused by cycles in the network.  These cycles would induce an infinite loop in the [sankey plugin](https://github.com/d3/d3-plugins/tree/master/sankey).  Fortunately, [Stefaan Lippens \@soxofaan](https://github.com/soxofaan) fixed and improved Sankey with [d3-plugin-captain-sankey]( https://github.com/soxofaan/d3-plugin-captain-sankey).  We were able to just plug it right in.  **I screwed up the CRAN `sankeyNetwork`, so to get the same output as below please install from Github**.

```{r "example02", eval = FALSE}
library(networkD3)

# simple network with cycles
net_cycles <- list(
  links = data.frame(
    source = c(0,0,0,1,1,5),
    target = c(1,2,3,4,5,0),
    value = 10
  ),
  nodes = data.frame(
    name = letters[1:6]
  )
)

# notice how few arguments we need now
# some output but not the nice output I expect
sankeyNetwork(
  net_cycles$links,
  net_cycles$nodes,
  Value = "value"
)

```

```{r "example02_run", eval = TRUE, echo = FALSE, fig.keep='none'}
library(networkD3)
library(jsonlite)

# simple network with cycles
net_cycles <- list(
  links = data.frame(
    source = c(0,0,0,1,1,5),
    target = c(1,2,3,4,5,0),
    value = 10
  ),
  nodes = data.frame(
    name = letters[1:6]
  )
)

# some output but not the nice output I expect
sankeyNetwork(
  net_cycles$links,
  net_cycles$nodes,
  Value = "value",
  width=400,height=400
) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example02.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 550
      , height = 550
      , overflow = "hidden"
   ))
```


## Thanks

Thanks Christopher Gandrud for spearheading `networkD3`.  Thanks Mike Bostock and Jason Davies for what seems like everything, but most especially `d3.js`.

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

