---
title: "Week32 - mapshaperWidget"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: ~
    toc: true
    mathjax: null
---

```{r echo=F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
)

library(pipeR)

```

## This Week's Widget - `mapshaperWidget`

Even a geo-novice like me can quickly see the amazing power of [`mapshaper`](https://github.com/mbloch/mapshaper) and it [web gui](http://www.mapshaper.org) and even use it immediately.  The author has clearly focused on a speedy, intuitive, but still full-feautured user experience.  To take it one step further though, let's try to integrate mapshaper seamlessly into our `R` geospatial workflow.

For now, I chose the awkward name `mapshaperWidget`, since "mapshaperr" is a little hard and there already is [`rmapshaper`](https://github.com/ateucher/rmapshaper) to use the non-gui node.js version.  I'm thinking if ok [with the author](https://github.com/ateucher/rmapshaper/issues/1), eventually I can fold `mapshaperWidget` into `rmapshaper`.

`mapshaper` is an entire website, so converting it to an `htmlwidget` is much different from most of the other conversions.  As of now, `mapshaperWidget` is not composeable and will not integrate into `rmarkdown`, since it completely takes over the page.  In later iterations, I hope to contain it.

`mapshaperWidget` relies on [`geojsonio`](https://github.com/ropensci/geojsonio).  Thanks to the [contributors](https://github.com/ropensci/geojsonio/graphs/contributors) for this very helpful package.  I also enjoyed adding some lazy magic from Hadley Wickham's [`lazyeval`](https://github.com/hadley/lazyeval).  Also, one last thanks to RStudio for integrating RStudio Viewer, which allows for a smooth all-in-one window experience.

## Quick Installation

`mapshaperWidget` is not yet on CRAN, so for now please install with `devtools::install_github`.

```{r eval = F}
devtools::install_github("timelyportfolio/mapshaper_htmlwidget")
```

## Examples

Examples usually help more than text.  Let's see if this is true.

### Straight `geojson`

```{r "example01", eval = FALSE}
# devtools::install_github("timelyportfolio/mapshaperWidget")
library(mapshaperWidget)

alabama <- paste0(
    readLines(
        "https://raw.githubusercontent.com/glynnbird/usstatesgeojson/master/alabama.geojson"
    )
    ,collapse = " "
)
mapshaper(alabama)
```

```{r "example01_run", eval = TRUE, echo = FALSE}
# devtools::install_github("timelyportfolio/mapshaperWidget")
library(mapshaperWidget)

alabama <- paste0(
    readLines(
        curl::curl("https://raw.githubusercontent.com/glynnbird/usstatesgeojson/master/alabama.geojson"
    ))
    ,collapse = " "
)
mapshaper(alabama,width=530,height=500) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example01.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 700
      , height = 670
      , overflow = "hidden"
   ))
```

### World Data from tmap

`data(World)` from [`tmap`](https://github.com/mtennekes/tmap) will give us a good example of working with `SpatialPolygonsDataFrame` in `mapshaper`.  Play with it a bit, and see if you can filter from the console.

```{r "example02", eval = FALSE}
# devtools::install_github("timelyportfolio/mapshaperWidget")
library(mapshaperWidget)
library(tmap)

data(World)
mapshaper(World)
```

```{r "example02_run", eval = TRUE, echo = FALSE}
# devtools::install_github("timelyportfolio/mapshaperWidget")
library(mapshaperWidget)
library(tmap)

data(World)
mapshaper(World,width=530,height=500) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example02.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 700
      , height = 670
      , overflow = "hidden"
   ))
```

### Multiple Spatial Objects

Now, let's incorporate some examples from `spdep` to show how we can analyze multiple spatial objects with `mapshaper`.  Remember, you can always drag and drop shapefiles, geojson, topojson files onto your `mapshaper`.  I highly recommend trying it out with the shapefiles from `spdep`.


```{r "example03", eval = FALSE}
# devtools::install_github("timelyportfolio/mapshaperWidget")
library(mapshaperWidget)
library(spdep)
library(maptools)

auckland <- readShapePoly(
  system.file(
    "etc/shapes/auckland.shp"
    ,package="spdep"
  )[1]
)
wheat <- readShapePoly(
  system.file(
    "etc/shapes/wheat.shp"
    ,package="spdep"
  )[1]
)
columbus <- readShapePoly(
  system.file(
    "etc/shapes/columbus.shp"
    ,package="spdep"
  )[1]
)

mapshaper( auckland, wheat, columbus )
```

```{r "example03_run", eval = TRUE, echo = FALSE}
# devtools::install_github("timelyportfolio/mapshaperWidget")
library(mapshaperWidget)
library(spdep)
library(maptools)

auckland <- readShapePoly(
  system.file(
    "etc/shapes/auckland.shp"
    ,package="spdep"
  )[1]
)
wheat <- readShapePoly(
  system.file(
    "etc/shapes/wheat.shp"
    ,package="spdep"
  )[1]
)
columbus <- readShapePoly(
  system.file(
    "etc/shapes/columbus.shp"
    ,package="spdep"
  )[1]
)

mapshaper( auckland, wheat, columbus,width=530,height=500) %>>%
  (
   htmlwidgets::as.iframe(
      .
      ,file="example03.html"
      , libdir = "lib"
      , selfcontained=F
      , width = 700
      , height = 670
      , overflow = "hidden"
   ))
```


## Next...

This is certainly not all, but I'll stop here. Next step will be bidirectional communication with `Shiny` and as said earlier some containment for composeability.  As always, I would love to work together if you have some ideas.


## Thanks

Thanks so much to [Matthew Bloch](https://github.com/mbloch) from the New York Times for `mapshaper`.  It is really good!

As always, thanks to 

- Ramnath Vaidyanathan and RStudio for `htmlwidgets`
- all the contributors to `R` and `JavaScript`

